### 2022-11-19 CBE_Sentiment.Rmd
Initial Wildfire Analysis

---
title: "Coffee Break Experiment #2"
author: "Sam Kalman and Charles Hutchinson"
date: "2022-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = F}
# Load Package
library(syuzhet)
library(tidyverse)
library(SentimentAnalysis)
library(ggridges)
```

```{r}
# Load and Join Bay Area Fire Tweets

# 4680 Tweets by 7 Variables
bay_area_tweets <- readRDS("bay_area_tweets.RDS")
# 3807 Users by 8 Variables
bay_area_users <- readRDS("bay_area_users.RDS")
colnames(bay_area_users)[1] <- "author_id"

# 3294 Unique Users
bay_area_users <- bay_area_users %>% 
  group_by(author_id) %>% 
  summarize_all(max)

# Join into one table
bay <- inner_join(bay_area_tweets, bay_area_users, by = "author_id") %>%
  mutate(fire = "Bay Area Fire")
```

```{r}
# Load and Join Camp Fire Tweets

# 50469 Tweets by 7 Variables
camp_tweets <- readRDS("camp_tweets.RDS")
# 44376 Users by 8 Variables
camp_users <- readRDS("camp_users.RDS")
colnames(camp_users)[1] <- "author_id"

# 26353 Unique Users
camp_users <- camp_users %>% 
  group_by(author_id) %>% 
  summarize_all(max)

# Join into one table
camp <- inner_join(camp_tweets, camp_users, by = "author_id") %>%
  mutate(fire = "Camp Fire")
```

```{r}
# Load and Join Dixie Fire Tweets

# 26139 Tweets by 7 Variables
dixie_tweets <- readRDS("dixie_tweets.RDS")
# 21230 Users by 8 Variables
dixie_users <- readRDS("dixie_users.RDS")
colnames(dixie_users)[1] <- "author_id"

# 11438 Unique Users
dixie_users <- dixie_users %>% 
  group_by(author_id) %>% 
  summarize_all(max)

# Join into one table
dixie <- inner_join(dixie_tweets, dixie_users, by = "author_id") %>%
  mutate(fire = "Dixie Fire")
```

```{r}
# Load and Join Tubbs Fire Tweets

# 3530 Tweets by 7 Variables
tubbs_tweets <- readRDS("tubbs_tweets.RDS")
# 2826 Users by 8 Variables
tubbs_users <- readRDS("tubbs_users.RDS")
colnames(tubbs_users)[1] <- "author_id"

# 2334 Unique Users
tubbs_users <- tubbs_users %>% 
  group_by(author_id) %>% 
  summarize_all(max)

# Join into one table
tubbs <- inner_join(tubbs_tweets, tubbs_users, by = "author_id") %>%
  mutate(fire = "Tubbs Fire")
```

```{r}
# Create Main Data Frame
main <- rbind(bay, camp, dixie, tubbs)
```

```{r}
# Pre-Processing
## Remove links and websites
main$text <- gsub("https.*","",main$text)
main$text <- gsub("http.*","",main$text)
main$text <- gsub("www.*","",main$text)
## Remove Twitter Handles
main$text <- gsub("@([a-zA-Z0-9]|[_])*","",main$text)

## Note: May need to add other pre-processing steps
```

```{r}
# Run Sentiment from syuzhet package
syuzhet_df <- get_sentiment(main$text) %>% as.data.frame()
```

```{r}
# Run Sentiment from SentimentAnalysis package
### Took 20ish minutes to run ###
analyzeSent_df <- analyzeSentiment(main$text)
```

```{r}
# Combine Sentiment Scores from both methods in main dataframe
main <- cbind(main, syuzhet_df, analyzeSent_df["SentimentQDAP"])
colnames(main)[16:17] <- c("syuzhet","SentimentAnalysis")
```

```{r}
# Save DataFrame as CSV
write.csv(main, "MainSentiment.csv", row.names = FALSE)
```

```{r}
# Syuzhet Range: -7.10 to 6.35
# SentimentAnalysis Range: -1 to 1
# > 0 is Positive for Both
# < 0 is Negative for Both
```

```{r}
# Add binary Positive and Negative values for both methods
main <- main %>% 
  mutate(syuzScore = case_when(syuzhet > 0 ~ "Positive", 
                               syuzhet < 0 ~ "Negative",
                               syuzhet == 0 ~ "Neutral"),
         saScore = case_when(SentimentAnalysis > 0 ~ "Positive", 
                             SentimentAnalysis < 0 ~ "Negative",
                             SentimentAnalysis == 0 ~ "Neutral"))

```

**Use Syuzhet Metric **

```{r}
 main %>%
  filter(followers_count > 0 & following_count > 0 & like_count > 0 & retweet_count > 0) %>%
  mutate(fire = fct_relevel(fire, c("Tubbs Fire", 
                      "Camp Fire", 
                      "Bay Area Fire", 
                      "Dixie Fire"))) %>%
  group_by(fire) %>%
  summarize(positive = length(syuzScore[syuzScore == "Positive"]) / n(),
            negative = length(syuzScore[syuzScore == "Negative"]) / n(),
            neutral = length(syuzScore[syuzScore == "Neutral"]) / n()) %>%
  gather(sentiment, percentage, positive:neutral) %>%
  ggplot(aes(x = fire, y = percentage, fill = sentiment)) +
  geom_bar(stat = "identity", position = "dodge") + 
  scale_fill_manual(values = c("red2", "blue", "darkgreen")) +
  theme_bw()
```

```{r}
# Remove News Organizations
 main %>%
  filter(followers_count > 0 & following_count > 0 & like_count > 0 & retweet_count > 0 & reply_count > 0) %>%
  filter(!stringr::str_detect(description,'news|meteorology')) %>%
  mutate(fire = fct_relevel(fire, c("Tubbs Fire", 
                      "Camp Fire", 
                      "Bay Area Fire", 
                      "Dixie Fire"))) %>%
  group_by(fire) %>%
  summarize(positive = length(syuzScore[syuzScore == "Positive"]) / n(),
            negative = length(syuzScore[syuzScore == "Negative"]) / n(),
            neutral = length(syuzScore[syuzScore == "Neutral"]) / n()) %>%
  gather(sentiment, percentage, positive:neutral) %>%
  ggplot(aes(x = fire, y = percentage, fill = sentiment)) +
  geom_bar(stat = "identity", position = "dodge") + 
  scale_fill_manual(values = c("red2", "blue", "darkgreen")) + 
  theme_bw()
```

```{r}
main %>%
  filter(followers_count > 0 & following_count > 0 & like_count > 0 & retweet_count > 0) %>%
  mutate(fire = fct_relevel(fire, c("Tubbs Fire", 
                      "Camp Fire", 
                      "Bay Area Fire", 
                      "Dixie Fire"))) %>%
  ggplot(aes(x = syuzhet)) +
  geom_density(stat = "density") +
  facet_wrap(~fire, ncol = 1) +
  theme_bw()
```

**Use SentimentAnalysis Metric**

```{r}
 main %>%
  mutate(fire = fct_relevel(fire, c("Tubbs Fire", 
                      "Camp Fire", 
                      "Bay Area Fire", 
                      "Dixie Fire"))) %>%
  group_by(fire) %>%
  summarize(positive = length(saScore[saScore == "Positive"]) / n(),
            negative = length(saScore[saScore == "Negative"]) / n(),
            neutral = length(saScore[saScore == "Neutral"]) / n()) %>%
  gather(sentiment, percentage, positive:neutral) %>%
  ggplot(aes(x = fire, y = percentage, fill = sentiment)) +
  geom_bar(stat = "identity", position = "dodge") + 
  theme_bw()
```

```{r}
# Remove News Organizations
 main %>%
  filter(!stringr::str_detect(description,'news|meteorology')) %>%
  mutate(fire = fct_relevel(fire, c("Tubbs Fire", 
                      "Camp Fire", 
                      "Bay Area Fire", 
                      "Dixie Fire"))) %>%
  group_by(fire) %>%
  summarize(positive = length(saScore[saScore == "Positive"]) / n(),
            negative = length(saScore[saScore == "Negative"]) / n(),
            neutral = length(saScore[saScore == "Neutral"]) / n()) %>%
  gather(sentiment, percentage, positive:neutral) %>%
  ggplot(aes(x = fire, y = percentage, fill = sentiment)) +
  geom_bar(stat = "identity", position = "dodge") + 
  theme_bw()
```

```{r}
main %>%
  mutate(fire = fct_relevel(fire, c("Tubbs Fire", 
                      "Camp Fire", 
                      "Bay Area Fire", 
                      "Dixie Fire"))) %>%
  ggplot(aes(x = SentimentAnalysis)) +
  geom_density(stat = "density") +
  facet_wrap(~fire, ncol = 1) +
  theme_bw()
```

### 2022-11-20 cbe2.Rmd
cbe2 intro + data sections

---
title: 'Coffee Break Experiment #2'
author: "Charles Hutchinson, Sam Kalman"
date: "Due 2022-11-20"
output: pdf_document
---

#### Introduction

One of North America's most noticeable symptoms of climate change has been a dramatic increase in the volume and intensity of wildfires, particularly in the western regions. August and September were previously defined by clear skies and fair weather, but fire and smoke have become a recurrent factor in this area. According to the National Interagency Fire Center (NIFC), more acreage burned nationally in each of the last seven years than any single year but one in the first ten years this data was collected (1983-1993, ex. 1988).^[See https://www.nifc.gov/fire-information/statistics/wildfires.] We sought to analyze if and how public sentiment has shifted over time with respect to these cataclysmic events, with the hopes of gaining some insight into public perceptions of climate change.

#### Data

We analyzed public sentiment towards wildfires by pulling publicly available tweets through Twitter's API. We chose to pull tweets made around the time of four of the largest, most recent wildfires in California. These events were the 2021 Dixie Fire, the 2020 Bay Area Fires, the 2018 Camp Fire, and the 2017 Tubbs Fire.

It was difficult to find relevant data. Tweets that were well-tagged (e.g., including #DixieFire for the Dixie Fire) tended to be associated with news organizations, which were overwhelmingly neutral and rarely "organic" reactions. However, loosening the search parameters inevitably lowered the data quality (e.g., searching for "smoke" mostly returned tweets referencing marijuana usage). It was especially challenging to develop searches with consistent methodology across separate wildfires, while still getting useful data.

We found our best results using the following approach:

* Use the commonly accepted fire name and "fire" as search terms. For example, the search terms for the 2021 Dixie Fire would simply be "dixie fire".
* Limit the search results to English. Creating sentiment dictionaries in other languages was out of scope for this project.
* Remove retweets.
* Limit the tweets to a month time frame. Some fires lasted longer than this, so we picked the interval in which the smoke was most impactful.

The size of each fire's corpus was unfortunately still far from uniform, but we found this to be a reasonable compromise among a set of unappealing options.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)

data.frame(
  Fire = c("Tubbs Fire", "Camp Fire", "Bay Area Fires", "Dixie Fire"),
  Tweets = c(3530, 50469, 4680, 26139)
) %>%
  kableExtra::kbl(caption = "Relevant Tweets per Recent Wildfire.", booktabs = T, linesep = "", digits = 2) %>%
  kableExtra::kable_styling(latex_options = "HOLD_position") %>%
  kableExtra::kable_classic()
```

### 2022-11-20 sample_map_code.Rmd
cbe2 intro + data sections

---
title: "Lab 12"
output:
  pdf_document:
    fig_caption: yes
    number_sections: true
author: "My Name"
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy.opts=list(width.cutoff=70), # this last bit auto-wraps code and comments so the don't run off the page, but you need to have formatR installed
  tidy=TRUE
)
```

# Lab 11: Geospacial Mapping: POP? SODA? or COKE? {-}

In 2013 a statistics (!) grad student at NC State named Joshua Katz put up a series of maps on an R Studio server that plotted the geographic distributions of lexical variants across the US. These turned out to be wildly popular and he published versions at Business Insider:

<https://www.businessinsider.com/american-english-dialects-maps-2018-1>

And at the New York Times, where he now works:

<https://www.nytimes.com/interactive/2014/upshot/dialect-quiz-map.html>

His maps are based on data collect by Bert Vaux and Scott Golder for the Harvard Dialect Survey in 2003:

<http://www4.uwm.edu/FLL/linguistics/dialect/index.html>

Katz's work is an interesting application of KNN to geospatial data. There is a nice introduction to the KNN algorithm here:

<https://www.analyticsvidhya.com/blog/2018/03/introduction-k-neighbours-algorithm-clustering/>

We're going to re-create Katz's pop vs. soda plot. This exercise requires a number of packages.

```{r setup, message = FALSE, error=FALSE, warning=FALSE}
# A couple for data wrangling...
library(tidyverse)
library(data.table)

# Our plots will be based on the results of knn...
library(kknn) # for knn

# We have some relatively intensive processing tasks that can be sped up... 
library(foreach) # for iterating over elements in a collection
library(doParallel) # for parallel processing

# And we'll be working with spatial data, so we need some specialized packages...
library(sf) # for spatial data
library(rnaturalearth) # map data for plotting
library(housingData) # geolocation data
require(rgeos)
```

# Prepare the Data

## Load the pop vs. soda data

The data can be downloaded from the **Data** folder on Canvas.

```{r warning=FALSE, message=FALSE}
pvs_df <- read_csv("/Users/user/Downloads/pop_vs_soda.csv", show_col_types = F)
```

Note how the data is structures.

```{r echo=FALSE}
pvs_df[,1:10] %>% head(10) %>% knitr::kable(format="latex", booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options=c("scale_down", "HOLD_position"))
```

[...]

```{r echo=FALSE}
pvs_df[,11:20] %>% head(10) %>%  knitr::kable(format="latex", booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options=c("scale_down", "HOLD_position"))
```

## Convert FIPS codes to longitude and latitude

We have counts by state and county. Importantly, those counts are identified by Federal Information Processing System (FIPS) codes. To convert that information to something mappable, we need GIS data for those FIPS codes. For that we can use data from the housingData package.

```{r}
hd <- housingData::geoCounty
hd$fips <- as.character(hd$fips)
```

Now let's select the columns we need from our data, and put that data in a long format.

```{r}
pvs_df <- pvs_df %>% select(fips = FIPS_combo, POP = SUMPOP, SODA = SUMSODA, 
                       COKE = SUMCOKE, OTHER = SUMOTHER)

pvs_df <- pvs_df %>% 
  pivot_longer(!fips, names_to = "var", values_to = "freq") %>% 
  filter(freq != 0) 
```

```{r echo=FALSE}
pvs_df %>% arrange(fips) %>% head(10) %>% knitr::kable(booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options="HOLD_position")
```

For the task, we need to convert counts into instances. For example, if there are 3 *soda* users in a county, we want
*soda*, *soda*, *soda* in the variable column. For that, we can use the data.table package.

```{r}
setDT(pvs_df)
pvs_df <- pvs_df[rep(seq(.N), freq), !"freq"]
```

```{r echo=FALSE}
pvs_df %>% arrange(fips) %>% head(10) %>% knitr::kable(booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options="HOLD_position")
```

Now let's join this with the GIS data to add longitudes and latitudes, and drop any empty rows.

```{r}
pvs_df <- left_join(pvs_df, select(hd, fips, lon, lat), by = "fips")
pvs_df <- pvs_df[!is.na(pvs_df$lon), ]
```

```{r echo=FALSE}
pvs_df %>% arrange(fips) %>% head(10) %>% knitr::kable(booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options="HOLD_position")
```

## Prepare a map

Now let's get the information we need for our map. We'll start with a map of the US states.

```{r}
states_us <- rnaturalearthdata::states50
```

There are several map options. We're going to select *iso_a2*. We'll also drop Alaska and Hawaii, so we'll only plot the Continental US.

```{r}
states_us <- states_us[states_us$iso_a2 == 'US',]
states_us <- states_us[ !grepl( "Alaska|Hawaii" , states_us$name ) , ]
```

Finally we want to convert this into an sf object for later plotting. This creates a *simple features* object. For more about *simple features* see here:

<https://r-spatial.github.io/sf/articles/sf1.html>

```{r}
states_us <- st_as_sf(states_us)
```

## Transform the data for plotting

Now we'll similarly convert the coordinates in our pop vs. soda data. Note that we point the *st_as_sf()* function to our "lon" and "lat" columns. Additionally, we need to specify the coordinate reference system (crs). This can be tricky... See here for a discussion:

<https://ryanpeek.github.io/2017-08-03-converting-XY-data-with-sf-package/>

We're specifying The World Geodetic System 1984 or WGS84.

```{r}
point_data <- pvs_df %>% st_as_sf(coords = c("lon", "lat"), crs = "+proj=longlat +ellps=WGS84")
```

Finally, we need to specify a projection. Remember we are "projecting" geographical space into 2 dimensions. So we have to decide what convention to use. We'll use the North America Lambert Conformal Conic projection:

<https://en.wikipedia.org/wiki/Lambert_conformal_conic_projection>

```{r}
nalcc <- "+proj=lcc +lat_1=20 +lat_2=60 +lat_0=40 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
```

Now, we can transform our US data to fit this projection. And likewise our pop vs. soda data.

```{r}
us <- states_us %>% st_transform(nalcc)
point_data <- point_data %>% st_transform(nalcc)
```

Next, we'll make a data.frame with 3 columns. And we'll convert our variable to a factor.

```{r}
dialects_train <- data.frame(dialect = point_data$var, 
                             lon = st_coordinates(point_data)[, 1], 
                             lat = st_coordinates(point_data)[, 2]) %>%
  mutate(dialect = as.factor(dialect))
```

```{r echo=FALSE}
dialects_train %>% head(10) %>% knitr::kable(booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options="HOLD_position")
```

To get a sense of our data, we can take a random sample from our training data and plot it.

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=7}
dialects_sample <- dialects_train %>% sample_n(1000)

ggplot(data = dialects_sample) +
  geom_sf(data = us) +
  geom_point(aes(x = lon, y = lat, color = dialect), alpha = 1) +
  scale_color_brewer(palette = "Dark2",  name = "Dialect")
```

# KNN classification

## Case study: Allegheny and Philadelphia counties

The implementation of the knn classifier is a little different from earlier classification problems, so let's walk through the basic idea with a quick example. First, let's subset out the data for 2 counties: Allegheny and Philadelphia counties (identified by their FIPS codes).

Then, we'll construct a simple data frame with 3 columns.

```{r}
dialects_penn <- point_data[grepl("42003|42101", point_data$fips), ]

dialects_penn <- data.frame(dialect = dialects_penn$var, 
                            lon = st_coordinates(dialects_penn)[, 1], 
                            lat = st_coordinates(dialects_penn)[, 2]) %>%
 mutate(dialect = as.factor(dialect))
```

From that, we can split the data into a training and test set.

```{r}
set.seed(123)
valid_split <- rsample::initial_split(dialects_penn, .75)

penn_train <- rsample::analysis(valid_split)
penn_test <- rsample::assessment(valid_split)
```

For more conventional classification tasks, we would want to determine the optimal k for our model using the **train.kknn()** function:

<https://rpubs.com/bonibruno/svm-knn>

For our purposes, we'll use a quick rule of thumb and use the square-root of our *n* observations. Now we can call a model in which we try to predict the dialect variant (*soda*, *pop*, *coke*, *other*) based on geolocation.

```{r}
knn_penn <- kknn::kknn(dialect ~ ., 
                       train = penn_train, 
                       test = penn_test, 
                       kernel = "gaussian", 
                       k = 61)
```

For conventional classification tasks, we would be interested in our confusion matrix, and our model's accuracy

```{r}
tab <- table(knn_penn$fitted.values, penn_test$dialect)

accuracy <- function(x){sum(diag(x)/(sum(rowSums(x)))) * 100}
accuracy(tab) %>% knitr::kable(booktabs = TRUE, digits = 2, col.names = "Accuracy")
```

For our purposes, we're interested in the probabilities of each variant being preferred based on location. Allegheny county is at the top of our table, with a very strong preference for *pop*.

```{r}
knn_penn$prob %>% head(1) %>% knitr::kable(booktabs = TRUE, digits = 2)
```

And Philadelphia county is at the bottom of our table with a strong preference for *soda*.

```{r}
knn_penn$prob %>% tail(1) %>% knitr::kable(booktabs = TRUE, digits = 2)
```

## Spatial interpolation

Those probabilities can be used to select both color (for the most probable variant) and the alpha of the color (based on the value of the probability).

Sounds easy enough, but we don't have values for the entire map. We need to use our known values to estimate those at other unknown points. This is process called spatial interpolation:

<https://docs.qgis.org/2.18/en/docs/gentle_gis_introduction/spatial_analysis_interpolation.html>

What follows in wholely indebted to Timo Grossenbacher.

<https://timogrossenbacher.ch/2018/03/categorical-spatial-interpolation-with-r/>

Essentially, we need to configure an empty grid and interpolate that grid with our point data. Another complication is that this process is quite memory and computationally intensive. So we're ultimately going to split up that grid into batches.

We'll begin setting up our grid by specifying its width in pixels.

```{r}
width_in_pixels <- 300
```

The width of a grid cell is calculated from the dimensions of our US map and is defined as dx.

```{r}
dx <- ceiling( (st_bbox(us)["xmax"] - st_bbox(us)["xmin"]) / width_in_pixels)
```

The height of a grid cell is the same because we'll use quadratic grid cells, **dx == dy**.

```{r}
dy <- dx
```

Calculate the height in pixels of the resulting grid.

```{r}
height_in_pixels <- floor( (st_bbox(us)["ymax"] - st_bbox(us)["ymin"]) / dy)
```

And finally construct the grid with the **st_make_grid()** function.

```{r}
grid <- st_make_grid(us, cellsize = dx, n = c(width_in_pixels, height_in_pixels), what = "centers")
```

Next, we'll set our k at 1000. This will result is a smoother looking map.

```{r}
k <- 1000
```

Now, we can define a **compute_grid()** function that returns the probability of the most likely variant.

```{r}
compute_grid <- function(grid, dialects_train, knn) {
  # create empty result data frame
  dialects_result <- data.frame(dialect = as.factor(NA), 
                                lon = st_coordinates(grid)[, 1], 
                                lat = st_coordinates(grid)[, 2])
  # run KKNN
  dialects_kknn <- kknn::kknn(dialect ~ ., 
                              train = dialects_train, 
                              test = dialects_result, 
                              kernel = "gaussian", 
                              k = knn)
  # bring back to result data frame
  # only retain the probability of the dominant dialect at that grid cell
  dialects_result <- dialects_result %>%
    # extract the interpolated dialect at each grid cell with the 
    # kknn::fitted() function
    mutate(dialect = fitted(dialects_kknn),
           # only retain the probability of the interpolated dialect,
           # discard the others
           prob = apply(dialects_kknn$prob, 1, function(x) max(x)))

  return(dialects_result)
}
```

We can set the parameters for parallel and batch processing. Adjust cores and batch size accordingly.

```{r}
registerDoParallel(cores = 4) # if you have more processing power you can up this number

# Specify number of batches and resulting size of each batch (in grid cells).
no_batches <- 60 # increase this number if you run into memory problems
batch_size <- ceiling(length(grid) / no_batches)
```

Finally, we'll generate our result using parallel processing. This will take at least a couple of minutes.

```{r}
dialects_result <- foreach(.packages = c("sf", "tidyverse"),
  batch_no = 1:no_batches, 
  # after each grid section is computed, rbind the resulting df into one big dialects_result df
  .combine = rbind, 
  # the order of grid computation doesn't matter: this speeds it up even more
  .inorder = FALSE) %dopar% {
    # compute indices for each grid section, depending on batch_size and current batch
    start_idx <- (batch_no - 1) * batch_size + 1
    end_idx <- batch_no * batch_size
    # specify which section of the grid to interpolate, using regular subsetting
    grid_batch <- grid[start_idx:ifelse(end_idx > length(grid), 
                                        length(grid),
                                        end_idx)]
    # apply the actual computation to each grid section
    df <- compute_grid(grid_batch, dialects_train, k)
  }
```

We'll now convert the result into an **sf** object for mapping.

```{r}
dialects_raster <- st_as_sf(dialects_result, 
                            coords = c("lon", "lat"),
                            crs = nalcc,
                            remove = F)
```

Remember that we've calculated a rectangular grid. As a last step, we can use the outline of the US to mask the borders of our map.

```{r}
dialects_raster <- dialects_raster[us, ]
```

And plot the result... (Note that we're dropping *other*.)

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=7}
ggplot(data = dialects_raster %>% filter(dialect != "OTHER")) +
  geom_raster(aes(x = lon, y = lat, fill = dialect, alpha = prob)) +
  scale_fill_manual(values = c("forestgreen", "steelblue", "tomato")) +
  scale_alpha(guide = 'none') +
  geom_sf(data = us, alpha = 0, size = 0.25) +
  theme_void() +
  theme(legend.position = c(0.1, 0.2), legend.box = "horizontal") +
  theme(legend.title=element_blank())
```

How does our result compare to Katz's?

<https://www.foodrepublic.com/2013/07/29/a-final-word-on-the-soda-vs-pop-vs-coke-nomenclature-debate/>

### 2022-11-20 load.R

@@ -5,7 +5,7 @@ library(tidyverse)
 # https://developer.twitter.com/en/docs/tutorials/getting-historical-tweets-using-the-full-archive-search-endpoint
 # https://developer.twitter.com/en/docs/twitter-api/tweets/search/integrate/build-a-query
 # https://developer.twitter.com/en/docs/twitter-api/data-dictionary/introduction
-bearer_token = readLines('~/msp/wildfire_sentiment_analysis/bearer_token.txt')
+bearer_token = readLines('~/text analysis/wildfire_sentiment_analysis/bearer_token.txt')
 
 headers = c(
   `Authorization` = sprintf('Bearer %s', bearer_token)

### 2022-11-20 cbe2.Rmd
Updated Methods and Results Section

@@
 
-#### Introduction
+**Introduction**
 
@@
 
-#### Data
+**Data**
 
@@
 ```
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+main <- read_csv("MainSentiment.csv")
+
+main <- main %>% 
+  mutate(syuzScore = case_when(syuzhet > 0 ~ "Positive", 
+                               syuzhet < 0 ~ "Negative",
+                               syuzhet == 0 ~ "Neutral"))
+```
+
+**Methods**
+
+To better understand how perception of climate change has changed over time, a sentiment analysis algorithm from the syuzhet^[See https://www.rdocumentation.org/packages/syuzhet/versions/1.0.6] package was applied to the 84,818 tweets across the 4 major fires of interest.
+The syuzhet package returns sentiment scores in the degree of positively/negativity using a sentiment dictionary developed in the Nebraska Literacy Labs.
+All links and web addresses were removed from the tweets as well as twitter handles, indicating that the tweet was mentioning or replying to another user.
+Each tweet was given a sentiment score.
+Scores above 0 were considered positive, below 0 were negative, and equal to 0 were considered a neutral tweet.
+Any tweets from twitter accounts with the words "news" or "meteorology" in their biographies were filtered out.
+We found that most news organization tweets contained neutral, non-opinionated information about the respective wildfire.
+To control for twitter bots adding unnecessary noise to the data, the users needed to have at least 1 follower and the tweet must be interacted with (like/retweet/comment) at least once.
+
+**Results**
+
+In Figure 1 below, the sentiment classification breakdown is plotted for each of the 4 wildfires.
+To account for the tweet volume difference between the four fires, percentages of sentiment type is calculated based on the total number of tweets pertaining to a respective fire.
+The fire are in chronological order, beginning with Tubbs Fire and ending with the Dixie Fire.
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3, fig.width=6,fig.cap="Shift in Wildfire Sentiment Over Time."}
+main %>%
+  filter(!stringr::str_detect(description,'news|meteorology')) %>%
+  filter(followers_count > 0 & 
+           (like_count > 0 | retweet_count > 0 | quote_count > 0)) %>%
+  mutate(fire = fct_relevel(fire, c("Tubbs Fire", 
+                      "Camp Fire", 
+                      "Bay Area Fire", 
+                      "Dixie Fire"))) %>%
+  group_by(fire) %>%
+  summarize(Positive = length(syuzScore[syuzScore == "Positive"]) / n(),
+            Negative = length(syuzScore[syuzScore == "Negative"]) / n(),
+            Neutral = length(syuzScore[syuzScore == "Neutral"]) / n()) %>%
+  gather(sentiment, percentage, Positive:Neutral) %>%
+  ggplot(aes(x = fire, y = percentage, fill = sentiment)) +
+  geom_bar(stat = "identity", position = "dodge") + 
+  scale_fill_manual(values = c("red2", "blue", "darkgreen")) + 
+  theme_bw() +
+  labs(title = "Wildfire Sentiment Change Over Time",
+       x = "",
+       y = "Percentage",
+       fill = "Sentiment")
+```
+
+Even though the tweet volume is not consistent throughout time, figure 1 gives us preliminary evidence that there may a shift in tweet sentiment as we observe more and more of these natural disasters.
+Beginning with Camp Fire, the amount of negative tweets has been steadily increasing for each wildfire.
+
+**Discussion**

### 2022-11-20 cbe2.Rmd
semi final draft

--- 2022-11-20 cbe2.Rmd
@@
 ```
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+main <- read_csv("MainSentiment.csv")
+
+main <- main %>% 
+  mutate(syuzScore = case_when(syuzhet > 0 ~ "Positive", 
+                               syuzhet < 0 ~ "Negative",
+                               syuzhet == 0 ~ "Neutral"))
+```
+
+**Methods**
+
+To better understand how perception of climate change has changed over time, we used a sentiment analysis algorithm from the syuzhet^[See https://www.rdocumentation.org/packages/syuzhet/versions/1.0.6] package on our Twitter data.
+The syuzhet package returns sentiment scores on a spectrum of positively/negativity using a sentiment dictionary developed in the Nebraska Literacy Labs.
+We removed all links, web addresses, and twitter handles from the tweets.
+Sentiment scores above 0 were labeled positive, below 0 labeled negative, and equal to 0 labeled as neutral.
+Any tweets from Twitter accounts with the words "news" or "meteorology" in their biographies were filtered out.
+We found that most tweets from news organization contained neutral, non-opinionated information, which wasn't useful for our research question.
+In an attempt to filter out bots adding unnecessary noise to the data, we only used tweets from users with at least one follower, and the tweet itself needed at least one external interaction (a like/retweet/comment).
+
+**Results**
+
+In Figure 1, the sentiment classification breakdown is plotted for each of the four wildfires.
+To account for the tweet volume difference between the four fires, percentages of sentiment type are calculated based on the total number of tweets pertaining to each fire.
+The fires are in chronological order, beginning with Tubbs Fire and ending with the Dixie Fire.
+The error bars reflect a 95% confidence interval for each proportion.
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3, fig.width=6,fig.cap="Shift in Wildfire Sentiment Over Time."}
+main %>%
+  filter(!stringr::str_detect(description,'news|meteorology')) %>%
+  filter(followers_count > 0 & 
+           (like_count > 0 | retweet_count > 0 | quote_count > 0)) %>%
+  mutate(fire = fct_relevel(fire, c("Tubbs Fire", 
+                      "Camp Fire", 
+                      "Bay Area Fire", 
+                      "Dixie Fire"))) %>%
+  group_by(fire) %>%
+  summarize(Positive = length(syuzScore[syuzScore == "Positive"]) / n(),
+            Negative = length(syuzScore[syuzScore == "Negative"]) / n(),
+            Neutral = length(syuzScore[syuzScore == "Neutral"]) / n(),
+            Total = n()) %>%
+  gather(sentiment, percentage, Positive:Neutral) %>%
+  mutate(
+    se = sqrt(percentage * (1 - percentage) / Total),
+    lower = percentage - 2 * se,
+    upper = percentage + 2 * se
+  ) %>%
+  ggplot(aes(x = fire, y = percentage, fill = sentiment)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  geom_errorbar(aes(ymin = lower, ymax = upper), position = "dodge", color = "black") +
+  scale_fill_manual(values = c("red2", "blue", "darkgreen")) + 
+  scale_y_continuous(labels=scales::percent) +
+  theme_bw() +
+  labs(title = "Wildfire Sentiment Change Over Time",
+       x = "",
+       y = "Percentage",
+       fill = "Sentiment") +
+  theme(legend.position = "bottom")
+```
+
+**Discussion**
+
+Even though the tweet volume is not consistent throughout time, Figure 1 gives us preliminary evidence that there may be a shift in tweet sentiment as we observe more and more of these natural disasters.
+Beginning with Camp Fire, the amount of negative tweets significantly increased for each wildfire.
+
+As previously mentioned, data quality was the limiting factor on this analysis.
+It was difficult to find a set of relevant tweets using a consistent methodology for each fire.
+Finding a balance between organic and relevant reactions ended up requiring more judgment than we would have liked.
+
+It was surprising that we could not find more tweets for the Bay Area and Tubbs fires, given their proximity to major urban centers.
+We originally suspected that the Camp Fire tweet set contained many false positives, but it appears that fire generated more engagement than expected due to the almost total devastation (and subsequent press coverage) of the towns of Paradise, Concow, Magalia, and Butte Creek Canyon.
+The Dixie Fire was similarly destructive.
+As the largest ever single wildfire in the state's history, part of its impact included popular Lassen National Park, which clearly increased engagement on Twitter.
+
+Our next area of research will be studying location-based elements of the wildfire tweets.
+Data quality issues were compounded in our initial attempts to download this data, since only a subset of Twitter users enable this functionality.
+However, knowing the location of each tweet will allow us to better understand if these changes in sentiment were localized, or on a national level.

### 2022-11-20 cbe2.Rmd
recompiled pdf

@@
        y = "Percentage",
-       fill = "Sentiment") +
-  theme(legend.position = "bottom")
+       fill = "Sentiment")
 ```

### 2022-11-20 cbe2.Rmd
final draft

@@
 
-One of North America's most noticeable symptoms of climate change has been a dramatic increase in the volume and intensity of wildfires, particularly in the western regions.
+One of North America's most noticeable symptoms of climate change has been a dramatic increase in the volume and intensity of wildfires, particularly in its western regions.
 August and September were previously defined by clear skies and fair weather, but fire and smoke have become a recurrent factor in this area.
@@
 
-We analyzed public sentiment towards wildfires by pulling publicly available tweets through Twitter's API.
+We analyzed sentiment towards wildfires by pulling publicly available tweets through Twitter's API.
 We chose to pull tweets made around the time of four of the largest, most recent wildfires in California.
@@
 
-To better understand how perception of climate change has changed over time, we used a sentiment analysis algorithm from the syuzhet^[See https://www.rdocumentation.org/packages/syuzhet/versions/1.0.6] package on our Twitter data.
+We used a sentiment analysis algorithm from the syuzhet^[See https://www.rdocumentation.org/packages/syuzhet/versions/1.0.6] package to analyze our Twitter data.
 The syuzhet package returns sentiment scores on a spectrum of positively/negativity using a sentiment dictionary developed in the Nebraska Literacy Labs.
 We removed all links, web addresses, and twitter handles from the tweets.
-Sentiment scores above 0 were labeled positive, below 0 labeled negative, and equal to 0 labeled as neutral.
+Sentiment scores above 0, below 0, and equal to 0 were respectively labeled as positive, negative, and neutral.
 Any tweets from Twitter accounts with the words "news" or "meteorology" in their biographies were filtered out.
@@
 
-Even though the tweet volume is not consistent throughout time, Figure 1 gives us preliminary evidence that there may be a shift in tweet sentiment as we observe more and more of these natural disasters.
+Even though the tweet volume was not consistent throughout time, Figure 1 gives us preliminary evidence that there may have been a shift in tweet sentiment towards wildfires in the last five years.
 Beginning with Camp Fire, the amount of negative tweets significantly increased for each wildfire.
 
-As previously mentioned, data quality was the limiting factor on this analysis.
+As mentioned previously, data quality was the limiting factor for this analysis.
 It was difficult to find a set of relevant tweets using a consistent methodology for each fire.
@@
 The Dixie Fire was similarly destructive.
-As the largest ever single wildfire in the state's history, part of its impact included popular Lassen National Park, which clearly increased engagement on Twitter.
+As the largest ever single wildfire in the state's history, part of its impact included popular Lassen National Park, the coverage of which likely increased engagement on Twitter.
 
 Our next area of research will be studying location-based elements of the wildfire tweets.
-Data quality issues were compounded in our initial attempts to download this data, since only a subset of Twitter users enable this functionality.
-However, knowing the location of each tweet will allow us to better understand if these changes in sentiment were localized, or on a national level.
+Data quality issues were compounded in our initial attempts to collect this data, since only a subset of Twitter users enable this functionality.
+However, knowing the location of each tweet will allow us to better understand to which extent these changes in sentiment were localized.

### 2022-12-10 load.R
map implementation

@@ -5,13 +5,13 @@ library(tidyverse)
 # https://developer.twitter.com/en/docs/tutorials/getting-historical-tweets-using-the-full-archive-search-endpoint
 # https://developer.twitter.com/en/docs/twitter-api/tweets/search/integrate/build-a-query
 # https://developer.twitter.com/en/docs/twitter-api/data-dictionary/introduction
-bearer_token = readLines('~/text analysis/wildfire_sentiment_analysis/bearer_token.txt')
+bearer_token = readLines('~/msp/wildfire_sentiment_analysis/bearer_token.txt')
 
 headers = c(
   `Authorization` = sprintf('Bearer %s', bearer_token)
 )
 
-n <- 50000
+n <- 100000
 
 # search in the US 
 # place_country:US
@@ -22,9 +22,9 @@ n <- 50000
 # camp fire: November 2018
 # tubbs fire: October 2017
 params = list(
-  `query` = '#CampFire -is:retweet lang:en place_country:US',
-  `start_time` = '2018-11-01T00:00:00.00Z',
-  `end_time` = '2018-12-01T00:00:00.00Z',
+  `query` = 'fire -is:retweet lang:en place_country:US',
+  `start_time` = '2020-08-15T00:00:00.00Z',
+  `end_time` = '2020-09-15T00:00:00.00Z',
   `max_results` = '500',
   `tweet.fields` = 'created_at,lang,public_metrics',
   `expansions` = 'geo.place_id',
@@ -123,8 +123,8 @@ while (nrow(tweets) < n) {
   
 }
 
-saveRDS(tweets, 'tubbs_tweets.RDS')
-saveRDS(users, 'tubbs_users.RDS')
+saveRDS(tweets, 'bay_area_fire_tweets.RDS')
+saveRDS(places, 'bay_area_fire_places.RDS')
 
 # t3 <- readRDS('dixie_tweets.RDS')
 

### 2022-12-10 map.R

@@ -1,0 +1,147 @@
+library(tidyverse)
+library(kknn)
+
+library(foreach)
+library(doParallel)
+
+library(sf)
+library(rnaturalearth)
+library(housingData)
+require(rgeos)
+
+
+setwd("~/msp/wildfire_sentiment_analysis/data")
+
+tweets <- readRDS('dixie_fire_tweets.RDS')
+places <- readRDS('dixie_fire_places.RDS')
+sentiment <- read.csv('dixie_fire_sentiment.csv')
+
+sentiment$text_cleaned <- sentiment$text
+sentiment$text <- NULL
+places <- distinct(places)
+
+pattern <- '^([^\\s]+) ([^\\s]+) ([^\\s]+) ([^\\s]+)$'
+
+data <- cbind(tweets, sentiment) %>%
+  # filter(like_count > 0 & retweet_count > 0) %>%
+  filter(label != "neutral") %>%
+  mutate(label = str_to_title(label)) %>%
+  left_join(places, by = c('place_id' = 'id')) %>%
+  transmute(
+    sentiment = as.factor(label),
+    long = (as.numeric(gsub(pattern, '\\1', bbox)) + 
+              as.numeric(gsub(pattern, '\\3', bbox)))/2,
+    lat = (as.numeric(gsub(pattern, '\\2', bbox)) + 
+             as.numeric(gsub(pattern, '\\4', bbox)))/2
+  )
+
+
+
+states_us <- rnaturalearthdata::states50
+
+states_us <- states_us[states_us$iso_a2 == 'US',]
+states_us <- states_us[ !grepl( "Alaska|Hawaii" , states_us$name ) , ]
+
+states_us <- st_as_sf(states_us)
+
+nalcc <- "+proj=lcc +lat_1=20 +lat_2=60 +lat_0=40 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
+
+us <- states_us %>% st_transform(nalcc)
+
+coord_data <- data %>%
+  drop_na() %>%
+  # limit to tweets in the lower 48.
+  filter(long < -65, long > -125) %>%
+  filter(lat < 50, lat > 24) %>%
+  st_as_sf(coords = c("long", "lat"), crs = "+proj=longlat +ellps=WGS84") %>%
+  st_transform(nalcc)
+
+map_df <- data.frame(
+  sentiment = coord_data$sentiment, 
[…]

### 2022-12-11 paper.Rmd
rough draft + code cleanup

---
title: "INSERT"
author: "Charles Hutchinson, Sam Kalman"
date: "December 12, 2022"
output: pdf_document
header-includes:
- \usepackage{booktabs}
- \usepackage{subfig}
- \usepackage{placeins}
---

```{r, include=FALSE}
library(tidyverse)
library(sf)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## Introduction

One of North America's most noticeable symptoms of climate change has been a dramatic increase in the volume and intensity of wildfires, particularly in its western regions. August and September were previously defined by clear skies and fair weather, but fire and smoke have become a recurrent factor in this area. According to the National Interagency Fire Center (NIFC), more acreage burned nationally in each of the last seven years than any single year but one in the first ten years this data was collected (1983-1993, ex. 1988).^[See https://www.nifc.gov/fire-information/statistics/wildfires.] We sought to analyze if and how public sentiment has shifted over time with respect to these cataclysmic events, with the hopes of gaining some insight into public perceptions of climate change.

## Data

We analyzed sentiment towards wildfires by pulling publicly available tweets through Twitter's API. We chose to pull tweets made around the time of four of the largest, most recent wildfires in California. These events were the 2021 Dixie Fire, the 2020 Bay Area Fires, the 2018 Camp Fire, and the 2017 Tubbs Fire.

It was difficult to find relevant data. Tweets that were well-tagged (e.g., including #DixieFire for the Dixie Fire) tended to be associated with news organizations, which were overwhelmingly neutral and rarely "organic" reactions. However, loosening the search parameters inevitably lowered the data quality (e.g., searching for "smoke" returned a large sample of tweets referencing marijuana usage). It was especially challenging to develop searches with consistent methodology across separate wildfires, while still getting useful data.

We found our best results using the following approach:

* Use the commonly accepted fire name and "fire" as search terms. For example, the search terms for the 2021 Dixie Fire would simply be "dixie fire".
* Limit the search results to English. Creating sentiment dictionaries in other languages was out of scope for this project.
* Remove retweets.
* Limit the tweets to a month time frame. Some fires lasted longer than this, so we picked the interval in which the smoke was most impactful.

The size of each fire's corpus was unfortunately still far from uniform, possibly due to unequal media coverage, but we found this to be a reasonable compromise among a set of unappealing options. We found both Camp Fire and Dixie Fire generated far more content on Twitter under our listed constraints when compared to Tubbs Fire and the Bay Area Fires. Table 1 shows the volume of tweets per fire.

```{r}
data.frame(
  Fire = c("Tubbs Fire", "Camp Fire", "Bay Area Fires", "Dixie Fire"),
  Tweets = c(3530, 50469, 4680, 26139),
  Words = c(51187, 1021221, 129082, 577785)
) %>%
  kableExtra::kbl(caption = "Relevant Tweets per Recent Wildfire.", booktabs = T, linesep = "") %>%
  kableExtra::kable_styling(latex_options = "HOLD_position") %>%
  kableExtra::kable_classic()
```

Our twitter license included access to a range of metadata. We chose to analyze tweet level information (specifically Likes, Retweet, Quote Tweets, and Date), as well as certain user level information (specifically Followers, Following, Biography, and Twitter Name).

Filtering for tweets that explicitly mentioned the fire by name significantly limited the total number of results. This proved to be particularly debilitating when working on location-based analyses, since only a subset of that limited population reported their location while tweeting. To alleviate this problem, we chose to use much more generic search terms. For the purposes of this paper, we found the search term "fire" tended to return the most relevant results (as compared to similarly broad terms like "smoke" or "burn"), though this was not a systematic decision. Nonetheless, it is hard to argue that "fire" is not an appropriate search term to use when conducting sentiment analysis on wildfires.

We continued to use the same date windows for each fire. As before, we only pulled English tweets, and filtered out any retweets. We asked the Twitter API to filter for tweets in the United States, then manually filtered those tweets down to those made in the contiguous lower 48 with longitude and latitude filters. Twitter provides locations in a bounding box, not on a single point, so we assumed that each tweet was made in the center of the bounding box. This was somewhat problematic when dealing with state-size bounding boxes, but the vast majority of locations were given on the city level, with much smaller bounding boxes.

```{r}
data.frame(
  Fire = c('Dixie', 'Bay Area', 'Camp', 'Tubbs'),
  Tweets = c(36353, 71608, 84284, 76839),
  Words = c(749509, 1468934, 1746368, 1165880)
) %>%
  kableExtra::kbl(caption = "Location-Enabled Tweet Counts", booktabs=T, linesep = "") %>%
  kableExtra::kable_styling(latex_options = "HOLD_position") %>%
  kableExtra::kable_classic()
```

## Methods

We used two sentiment analysis algorithms of varying complexity to classify the aforementioned tweets. The first algorithm we used was from the `syuzhet` package^[https://www.rdocumentation.org/packages/syuzhet/versions/1.0.6] in R. The syuzhet package returns sentiment scores on a spectrum of positively/negativity using a sentiment dictionary developed in the Nebraska Literacy Labs. We deemed any sentiment score that was greater than zero to be positive, less than zero, negative, and exactly equal to zero, neutral. Unfortunately, the `syuzhet` package wasn't designed or trained on short, declarative statements such as tweets.

To overcome this issue, we decided to use Cardiff's roBERTa-base model for sentiment analysis.^[https://huggingface.co/cardiffnlp/twitter-roBERTa-sentiment-latest] This model was trained on roughly 124 million tweets from January 2018 to December 2021 and used the TweetEval benchmark for fine tuning the sentiment analysis. The algorithm outputs probabilities that a tweet is positive, neutral, or negative. For the purposes of our analysis, we simplified the output to a categorical variable representing the sentiment with the highest probability. (Strength of sentiment was not captured.)

BERT models are trained on both language modeling and next sentence prediction. It is important to note that BERT models generally do not use words as tokens - they use a combination of word-level, segment-level, and position-level tokens to encode the input text. Word-level tokens are individual words or sub-words in the input text, segment-level tokens are used to indicate whether a given word is part of the first or second sentence in a pair of sentences (in the case of sentence-pair tasks), and position-level tokens are used to indicate the position of each word within the input text. This may convey some of the complexity of the contextual embedding performed in the model. Regardless, we note that the word counts in Tables 1 and 2 are not accurate token counts for Cardiff's roBERTa model.

We found that most tweets from news organization contained neutral, non-opinionated information, which wasn’t useful for our research question. Therefore ,any tweet that came from a user with the word "meteorology" in their bio was not considered in our analysis. Additionally, in an attempt to filter out bots adding unnecessary noise to the data, we only used tweets from users with at least one follower, and the tweet itself needed at least one external interaction (a like/retweet/comment). This was done consistently before running both of the discussed sentiment algorithms. It's worth noting that the authors of Cardiff's roBERTa model (Loureriro et al., 2022) address the bot issue by filtering out the top 1% of users by volume, but we did not have a similarly unbiased data set to determine an appropriate posting cutoff.

To track to sentiment shift through time, we chronologically ordered the fires (Tubbs Fire, Camp Fire, Bay Area Fire, Dixie Fire) and calculated the proportion of each sentiment for each fire. We used proportions because of the unequal distribution of tweets across the fires.

We first converted from longitude/latitude coordinates to WGS84 coordinates, then projected these coordinates using a Lambert Conformal Conic projection (LCC). This process was completed using the `sf` package. [ADD MORE HERE]

## Results

In Figure 1, we plot the sentiment classification proportions for each of the four wildfires. The fires are in chronological order, beginning with Tubbs Fire and ending with the Dixie Fire. The error bars reflect a 95% confidence interval for each proportion.

```{r, fig.height=3, fig.width=6,fig.cap="Shift in Wildfire Sentiment Over Time."}

main <- read_csv("data/MainSentiment.csv")

main %>%
  filter(followers_count > 0 & 
           following_count > 0 & 
           like_count > 0) %>%
  filter(!stringr::str_detect(description,'meteorology')) %>%
  mutate(fire = fct_relevel(fire, c("Tubbs Fire", 
                      "Camp Fire", 
                      "Bay Area Fire", 
                      "Dixie Fire"))) %>%
  group_by(fire) %>%
  summarize(Positive = length(syuzhet_sent[syuzhet_sent == "positive"]) / n(),
            Negative = length(syuzhet_sent[syuzhet_sent == "negative"]) / n(),
            Neutral = length(syuzhet_sent[syuzhet_sent == "neutral"]) / n(),
            Total = n()) %>%
  gather(sentiment, percentage, Positive:Neutral) %>%
  mutate(
    se = sqrt(percentage * (1 - percentage) / Total),
    lower = percentage - 2 * se,
    upper = percentage + 2 * se) %>%
  ggplot(aes(x = fire, y = percentage, fill = sentiment)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.3) + 
  geom_errorbar(aes(ymin = lower, ymax = upper)
                , position = "dodge"
                , color = "black") +
  scale_fill_manual(values = c("red2", "blue", "darkgreen")) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Syuzhet Sentiment Proportions",
       x = "Fire",
       y = "Proportion of Tweets",
       fill = "Sentiment") +
  theme_bw()
```

There are not many neutral tweets, though our definition of neutrality with the `syuzhet` scoring is quite stringent. The majority of tweets are classified as negative, and the proportion of negativity appears to be increasing through time.

We repeat the same analysis using scores from Cardiff's roBERTa model in Figure 2.

```{r, fig.height=3, fig.width=6,fig.cap="Shift in Wildfire Sentiment Over Time."}
main %>%
  filter(followers_count > 0 & 
           following_count > 0 & 
           like_count > 0) %>%
  filter(!stringr::str_detect(description,'meteorology')) %>%
  mutate(fire = fct_relevel(fire, c("Tubbs Fire", 
                      "Camp Fire", 
                      "Bay Area Fire", 
                      "Dixie Fire"))) %>%
  group_by(fire) %>%
  summarize(Positive = length(bert_sent[bert_sent == "positive"]) / n(),
            Negative = length(bert_sent[bert_sent == "negative"]) / n(),
            Neutral = length(bert_sent[bert_sent == "neutral"]) / n(),
            Total = n()) %>%
  gather(sentiment, percentage, Positive:Neutral) %>%
  mutate(
    se = sqrt(percentage * (1 - percentage) / Total),
    lower = percentage - 2 * se,
    upper = percentage + 2 * se) %>%
  ggplot(aes(x = fire, y = percentage, fill = sentiment)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.3) + 
  geom_errorbar(aes(ymin = lower, ymax = upper)
                , position = "dodge"
                , color = "black") +
  scale_fill_manual(values = c("red2", "blue", "darkgreen")) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "roBERTa Sentiment Proportions",
       x = "Fire",
       y = "Proportion of Tweets",
       fill = "Sentiment") +
  theme_bw()

```

The roBERTa model classifies a much larger proportion of the tweets as neutral. In previous iterations of this analysis, it was difficult for us to remove all news related tweets. We found that this iteration was very effective at putting "inorganic" reactions from news organizations in this neutral category. However, the roBERTa model classified less negativity for the Tubbs fire than the syuzhet package and we see a more consistent increase in negativity, plateauing at the Bay Area Fire and Dixie Fire. It was difficult for us to remove all news related tweets, roBERTa model

Let's examine differences in specific tweet classifation between the syuzhet and roBERTa model. Specifically, for the Tubbs fire where we saw less negativity classified by the roBERTa model. It was difficult for us to remove all news related tweets. roBERTa model classified a lot of these as neutral when the syuzhet package would incorrectly assig the tweet as slightly positive or slightly negative.

**Examples of neutral for roBERTa and negative for syuzhet**

* Where Did the Tubbs Fire Begin? Investigators Scour the Ashes https://t.co/9z2QjBXi77

* The cause of the massive Tubbs fire is still unknown -- but officials think they may be zeroing in: https://t.co/0k9w4nFgpE.

These are examples of truly neutral tweets from news organizations. However, the syuzhet package tends to not assign many neutral tweets. These were incorrectly classified as negative by syuzhet, possibly due to words like "scour" or "unknown", but correctly seen as neutral from roBERTa.

**Examples of negative for roBERTa and positive for syuzhet**

* The personal treasures former Raiders great Cliff Branch lost in Santa Rosa fires--a truly sad tale by Bob Padecky. https://t.co/gIT0AulURq

* My parents lost everything in the Tubbs Fire yesterday. Anything will help. Please share as much as you can. Thank you so much. #tubbsfire https://t.co/xFAexYshSL

* Many Sonic staff in Santa Rosa and surrounding area are affected by the Tubbs Fire. Best wishes to all, and be safe! It is devastating. https://t.co/pVqAFisxKd

The first tweet used the word "great" to describe the football player but the overall sentiment should have been negative due to Cliff losing his possessions in the fire. Many tweets consisting of "Thank you" or "Be safe" were seen as positive from the syuzhet model, when the overall message of the tweet was in a negative context. This subtle distinction was correctly picked up by the roBERTa model.

**Examples of positive for roBERTa and negative for syuzhet**

* This Friday's midnight showing the Rocky Horror Picture Show is now a benefit for Tubbs Fire victims: https://t.co/Su7tOzcnfm

* All I can say is wow. Props to Berkeley #firefighters. Crazy video of firefighters stunned by scope of Tubbs Fire https://t.co/8kZwh6g5uy

These two examples highlight how the syuzhet package was tricked into a negative classification when the tweet should have been seen as positive. The describing of the movie has some words associated with negativity but the overall message of tweet was a benefit for Tubbs Fire victim, this was correctly picked up on by roBERTa. The second tweet is a user thanking the firefighters, the fact that she was stunned by the scope of the fire made syuzhet believe the tweet was negative. However, she tweeted this to give props to the firefighters. These examples highlight how the roBERTa model was more effective at distinguishing subtle nuances in tweets, resulting in more accurate classifications of sentiment.

Important to note that Tubbs was much shorter, and the other summer fires were usually coincident with a broader wildfire season. [ADD MORE HERE]

```{r, fig.cap = "Tweet Sentiment (roBERTa) by Location", fig.height = 7, fig.subcap = c("Dixie (2021)", "Bay Area (2020)", "Camp (2018)", "Tubbs (2017)"), fig.ncol = 2, out.width = "50%", fig.align='center'}

raster_df <- readRDS('data/dixie_map_data.RDS')
us <- readRDS('data/us_map_data.RDS')

ggplot(data = raster_df) +
  geom_raster(aes(x = lon, y = lat, fill = sentiment, alpha = prob)) +
  scale_fill_manual(values = c("tomato", "steelblue", "forestgreen")) +
  scale_alpha(guide = 'none') +
  geom_sf(data = us, alpha = 0, size = 0.25) +
  theme_void() +
  theme(legend.position = c(0.1, 0.2), legend.box = "horizontal") +
  theme(legend.title=element_blank())

raster_df <- readRDS('data/bay_area_map_data.RDS')

ggplot(data = raster_df) +
  geom_raster(aes(x = lon, y = lat, fill = sentiment, alpha = prob)) +
  scale_fill_manual(values = c("tomato", "steelblue", "forestgreen")) +
  scale_alpha(guide = 'none') +
  geom_sf(data = us, alpha = 0, size = 0.25) +
  theme_void() +
  theme(legend.position = c(0.1, 0.2), legend.box = "horizontal") +
  theme(legend.title=element_blank())

raster_df <- readRDS('data/camp_map_data.RDS')

ggplot(data = raster_df) +
  geom_raster(aes(x = lon, y = lat, fill = sentiment, alpha = prob)) +
  scale_fill_manual(values = c("tomato", "steelblue", "forestgreen")) +
  scale_alpha(guide = 'none') +
  geom_sf(data = us, alpha = 0, size = 0.25) +
  theme_void() +
  theme(legend.position = c(0.1, 0.2), legend.box = "horizontal") +
  theme(legend.title=element_blank())

raster_df <- readRDS('data/tubbs_map_data.RDS')

ggplot(data = raster_df) +
  geom_raster(aes(x = lon, y = lat, fill = sentiment, alpha = prob)) +
  scale_fill_manual(values = c("tomato", "steelblue", "forestgreen")) +
  scale_alpha(guide = 'none') +
  geom_sf(data = us, alpha = 0, size = 0.25) +
  theme_void() +
  theme(legend.position = c(0.1, 0.2), legend.box = "horizontal") +
  theme(legend.title=element_blank())
```

## Discussion

Even though the tweet volume was not consistent throughout time, Figure 1 gives us preliminary evidence that there may have been a shift in tweet sentiment towards wildfires in the last five years. Beginning with Camp Fire, the amount of negative tweets significantly increased for each wildfire.

As mentioned previously, data quality was the limiting factor for this analysis. It was difficult to find a set of relevant tweets using a consistent methodology for each fire. Finding a balance between organic and relevant reactions ended up requiring more judgment than we would have liked.

It was surprising that we could not find more tweets for the Bay Area and Tubbs fires, given their proximity to major urban centers. We originally suspected that the Camp Fire tweet set contained many false positives, but it appears that fire generated more engagement than expected due to the almost total devastation (and subsequent press coverage) of the towns of Paradise, Concow, Magalia, and Butte Creek Canyon. The Dixie Fire was similarly destructive. As the largest ever single wildfire in the state's history, part of its impact included popular Lassen National Park, the coverage of which likely increased engagement on Twitter.

### 2022-12-11 sentiment_eda.Rmd
rough draft + code cleanup

---
title: "Final Paper Part 1"
author: "Sam Kalman and Charles Hutchinson"
date: "2022-12-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = F}
# Load Package
library(syuzhet)
library(tidyverse)
```

```{r, message = F}
# Load and Join Bay Area Fire Tweets

# 4680 Tweets by 7 Variables
bay_area_tweets <- readRDS("bay_area_tweets.RDS")
# 3807 Users by 8 Variables
bay_area_users <- readRDS("bay_area_users.RDS")
colnames(bay_area_users)[1] <- "author_id"

bay_area_bert <- read_csv("bay_area_sentiment.csv")

bay_area_tweets <- cbind(bay_area_tweets, 
                         bay_area_bert$label, 
                         bay_area_bert$score)

colnames(bay_area_tweets)[8] <- "bert_sent"
colnames(bay_area_tweets)[9] <- "bert_score"

# 3294 Unique Users
bay_area_users <- bay_area_users %>% 
  group_by(author_id) %>% 
  summarize_all(max)

# Join into one table
bay <- inner_join(bay_area_tweets, bay_area_users, by = "author_id") %>%
  mutate(fire = "Bay Area Fire")
```

```{r, message = F}
# Load and Join Camp Fire Tweets

# 50469 Tweets by 7 Variables
camp_tweets <- readRDS("camp_tweets.RDS")
# 44376 Users by 8 Variables
camp_users <- readRDS("camp_users.RDS")
colnames(camp_users)[1] <- "author_id"

camp_bert <- read_csv("camp_sentiment.csv")

camp_tweets <- cbind(camp_tweets, 
                         camp_bert$label, 
                         camp_bert$score)

colnames(camp_tweets)[8] <- "bert_sent"
colnames(camp_tweets)[9] <- "bert_score"

# 26353 Unique Users
camp_users <- camp_users %>% 
  group_by(author_id) %>% 
  summarize_all(max)

# Join into one table
camp <- inner_join(camp_tweets, camp_users, by = "author_id") %>%
  mutate(fire = "Camp Fire")
```

```{r, message = F}
# Load and Join Dixie Fire Tweets

# 26139 Tweets by 7 Variables
dixie_tweets <- readRDS("dixie_tweets.RDS")
# 21230 Users by 8 Variables
dixie_users <- readRDS("dixie_users.RDS")
colnames(dixie_users)[1] <- "author_id"

dixie_bert <- read_csv("dixie_sentiment.csv")

dixie_tweets <- cbind(dixie_tweets, 
                         dixie_bert$label, 
                         dixie_bert$score)

colnames(dixie_tweets)[8] <- "bert_sent"
colnames(dixie_tweets)[9] <- "bert_score"

# 11438 Unique Users
dixie_users <- dixie_users %>% 
  group_by(author_id) %>% 
  summarize_all(max)

# Join into one table
dixie <- inner_join(dixie_tweets, dixie_users, by = "author_id") %>%
  mutate(fire = "Dixie Fire")
```

```{r, message = F}
# Load and Join Tubbs Fire Tweets

# 3530 Tweets by 7 Variables
tubbs_tweets <- readRDS("tubbs_tweets.RDS")
# 2826 Users by 8 Variables
tubbs_users <- readRDS("tubbs_users.RDS")
colnames(tubbs_users)[1] <- "author_id"

tubbs_bert <- read_csv("tubbs_sentiment.csv")

tubbs_tweets <- cbind(tubbs_tweets, 
                         tubbs_bert$label, 
                         tubbs_bert$score)

colnames(tubbs_tweets)[8] <- "bert_sent"
colnames(tubbs_tweets)[9] <- "bert_score"

# 2334 Unique Users
tubbs_users <- tubbs_users %>% 
  group_by(author_id) %>% 
  summarize_all(max)

# Join into one table
tubbs <- inner_join(tubbs_tweets, tubbs_users, by = "author_id") %>%
  mutate(fire = "Tubbs Fire")
```

```{r}
# Create Main Data Frame
main <- rbind(bay, camp, dixie, tubbs)

# Keep a copy of original tweet
main$orig_text <- main$text
```

```{r}
# Pre-Processing
## Remove links and websites
main$text <- gsub("https.*","",main$text)
main$text <- gsub("http.*","",main$text)
main$text <- gsub("www.*","",main$text)
## Remove Twitter Handles
main$text <- gsub("@([a-zA-Z0-9]|[_])*","",main$text)
```

```{r}
# Run Sentiment from syuzhet package for pre-processed text
syuzhet_df <- get_sentiment(main$text) %>% as.data.frame()
```

```{r}
# Add syuzhet sentiment score to main dataframe
main <- cbind(main, syuzhet_df)
colnames(main)[19] <- "syuzhet_score"
```

```{r}
# Syuzhet Range: -7.10 to 6.35
# > 0 is Positive
# < 0 is Negative
# = 0 is Neutral
# Add syuzhet sentiment score classification

main <- main %>% 
  mutate(syuzhet_sent = case_when(syuzhet_score > 0 ~ "positive", 
                               syuzhet_score < 0 ~ "negative",
                               syuzhet_score == 0 ~ "neutral"))
```

```{r}
# Save DataFrame as CSV
write.csv(main, "MainSentiment.csv", row.names = FALSE)
```

**Syuzhet**

```{r}
main %>%
  filter(followers_count > 0 & 
           following_count > 0 & 
           like_count > 0) %>%
  filter(!stringr::str_detect(description,'news')) %>%
  mutate(fire = fct_relevel(fire, c("Tubbs Fire", 
                      "Camp Fire", 
                      "Bay Area Fire", 
                      "Dixie Fire"))) %>%
  group_by(fire) %>%
  summarize(Positive = length(syuzhet_sent[syuzhet_sent == "positive"]) / n(),
            Negative = length(syuzhet_sent[syuzhet_sent == "negative"]) / n(),
            Neutral = length(syuzhet_sent[syuzhet_sent == "neutral"]) / n(),
            Total = n()) %>%
  gather(sentiment, percentage, Positive:Neutral) %>%
  mutate(
    se = sqrt(percentage * (1 - percentage) / Total),
    lower = percentage - 2 * se,
    upper = percentage + 2 * se) %>%
  ggplot(aes(x = fire, y = percentage, fill = sentiment)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.3) + 
  geom_errorbar(aes(ymin = lower, ymax = upper)
                , position = "dodge"
                , color = "black") +
  scale_fill_manual(values = c("red2", "blue", "darkgreen")) +
  labs(title = "Syuzhet Sentiment Analysis",
       x = "Fire",
       y = "Proportion of Tweets",
       fill = "Sentiment") +
  theme_bw()
```

**BERT**

```{r}
 main %>%
  filter(followers_count > 0 & 
           following_count > 0 & 
           like_count > 0) %>%
  filter(!stringr::str_detect(description,'news')) %>%
  mutate(fire = fct_relevel(fire, c("Tubbs Fire", 
                      "Camp Fire", 
                      "Bay Area Fire", 
                      "Dixie Fire"))) %>%
  group_by(fire) %>%
  summarize(Positive = length(bert_sent[bert_sent == "positive"]) / n(),
            Negative = length(bert_sent[bert_sent == "negative"]) / n(),
            Neutral = length(bert_sent[bert_sent == "neutral"]) / n(),
            Total = n()) %>%
  gather(sentiment, percentage, Positive:Neutral) %>%
  mutate(
    se = sqrt(percentage * (1 - percentage) / Total),
    lower = percentage - 2 * se,
    upper = percentage + 2 * se) %>%
  ggplot(aes(x = fire, y = percentage, fill = sentiment)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.3) + 
  geom_errorbar(aes(ymin = lower, ymax = upper)
                , position = "dodge"
                , color = "black") +
  scale_fill_manual(values = c("red2", "blue", "darkgreen")) +
  labs(title = "BERT Sentiment Analysis",
       x = "Fire",
       y = "Proportion of Tweets",
       fill = "Sentiment") +
  theme_bw()
```

### 2022-12-11 sk_draft.Rmd
rough draft + code cleanup

---
title: "668 Final Part 1"
author: "Sam Kalman and Charles Hutchinson"
date: "2022-12-11"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = F, include = F}
# Load Package
library(syuzhet)
library(tidyverse)
```

**Data**

We analyzed sentiment towards wildfires by pulling publicly available tweets through Twitter's API. We chose to pull tweets made around the time of four of the largest, most recent wildfires in California. These events were the 2021 Dixie Fire, the 2020 Bay Area Fires, the 2018 Camp Fire, and the 2017 Tubbs Fire.

It was difficult to find relevant data. Tweets that were well-tagged (e.g., including #DixieFire for the Dixie Fire) tended to be associated with news organizations, which were overwhelmingly neutral and rarely "organic" reactions. However, loosening the search parameters inevitably lowered the data quality (e.g., searching for "smoke" mostly returned tweets referencing marijuana usage). It was especially challenging to develop searches with consistent methodology across separate wildfires, while still getting useful data.

We found our best results using the following approach:

* Use the commonly accepted fire name and "fire" as search terms. For example, the search terms for the 2021 Dixie Fire would simply be "dixie fire".
* Limit the search results to English. Creating sentiment dictionaries in other languages was out of scope for this project.
* Remove retweets.
* Limit the tweets to a month time frame. Some fires lasted longer than this, so we picked the interval in which the smoke was most impactful.

The size of each fire's corpus was unfortunately still far from uniform, possibly due to unequal media coverage, but we found this to be a reasonable compromise among a set of unappealing options. We found both Camp Fire and Dixie Fire received far more media coverage, given our listed constraints, compared to Tubbs Fire and the Bay Area Fires. Below in Table 1 is the breakdown of tweets per fire.

```{r, echo = F, warning = F, message = F}
data.frame(
  Fire = c("Tubbs Fire", "Camp Fire", "Bay Area Fires", "Dixie Fire"),
  Tweets = c(3530, 50469, 4680, 26139)
) %>%
  kableExtra::kbl(caption = "Relevant Tweets per Recent Wildfire.", booktabs = T, linesep = "", digits = 2) %>%
  kableExtra::kable_styling(latex_options = "HOLD_position") %>%
  kableExtra::kable_classic()
```

Each tweet also included associated Meta Data. The Meta Data was broken into tweet level information (Likes, Retweet, Quote Tweets, Date) and user level information (Followers, Following, Biography, Twitter Name).

**Methods**

**Wildfire Sentiment Through Time**

In order to track the changes of wildfire Sentiment through time, we used two different sentiment analysis algorithms to get a sentiment classification for each tweet. Since sentiment analysis is an unsupervised modeling technique, meaning there isn't a correct sentiment to check each modeling prediction against, we decided on using two different algorithms and comparing the results across models. The first algorithm we used was from the syuzhet package^[https://www.rdocumentation.org/packages/syuzhet/versions/1.0.6] in R. The syuzhet package returns sentiment scores on a spectrum of positively/negativity using a sentiment dictionary developed in the Nebraska Literacy Labs. Any sentiment score that was greater than 0 was deem "Positive", less than 0, "Negative, and exactly equal to 0 was considered, "Neutral". However, a draw back of the syuzhet is that it is designed to highlight latent structure of narrative. This is accomplished by, "revealing emotional shifts that serve as proxies for narrative movement between conflict and conflict resolution". The syuzhet package wasn't designed or trained on short, declarative statements such as tweets.

To overcome this issue, we also decided to model sentiment through the cutting edge roBERTa-base model^[https://huggingface.co/cardiffnlp/twitter-roberta-base-sentiment-latest]. This model was trained on roughly 124 million tweets from January 2018 to December 2021 and used the TweetEval benchmark for fine tuning the sentiment analysis. Sentiment Analysis was given in three categories for the roBERTa-base model: degree of positivity, degree of negativity and degree in which the tweet was neutral. We used the largest score out of the three categories as the binary classification for the sentiment of a tweet.

We found that most tweets from news organization contained neutral, non-opinionated information, which wasn’t useful for our research question. Therefore any tweet that came from a user with the word "meterology" in their bio was not considered. Also, in an attempt to filter out bots adding unnecessary noise to the data, we only used tweets from users with at least one follower, and the tweet itself needed at least one external interaction (a like/retweet/comment). This was done consistently before running both of the discussed sentiment algorithms.

To track to sentiment shift through time, we chronologically ordered the fires (Tubbs Fire, Camp Fire, Bay Area Fire, Dixie Fire) and calculated the proportion of each sentiment for each fire. The fires represents a point in time and proportion was used because of the unequal distribution of tweets within each fire.

\pagebreak

**Results**

```{r, echo=FALSE, warning=FALSE, message=FALSE}
main <- read_csv("MainSentiment.csv") 
```

**Wildfire Sentiment Through Time**

In Figure 1, the sentiment classification breakdown is plotted for each of the four wildfires. To account for the tweet volume difference between the four fires, percentages of sentiment type are calculated based on the total number of tweets pertaining to each fire. The fires are in chronological order, beginning with Tubbs Fire and ending with the Dixie Fire. The error bars reflect a 95% confidence interval for each proportion.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3, fig.width=6,fig.cap="Shift in Wildfire Sentiment Over Time."}

main <- read_csv("MainSentiment.csv")

main %>%
  filter(followers_count > 0 & 
           following_count > 0 & 
           like_count > 0) %>%
  filter(!stringr::str_detect(description,'meteorology')) %>%
  mutate(fire = fct_relevel(fire, c("Tubbs Fire", 
                      "Camp Fire", 
                      "Bay Area Fire", 
                      "Dixie Fire"))) %>%
  group_by(fire) %>%
  summarize(Positive = length(syuzhet_sent[syuzhet_sent == "positive"]) / n(),
            Negative = length(syuzhet_sent[syuzhet_sent == "negative"]) / n(),
            Neutral = length(syuzhet_sent[syuzhet_sent == "neutral"]) / n(),
            Total = n()) %>%
  gather(sentiment, percentage, Positive:Neutral) %>%
  mutate(
    se = sqrt(percentage * (1 - percentage) / Total),
    lower = percentage - 2 * se,
    upper = percentage + 2 * se) %>%
  ggplot(aes(x = fire, y = percentage, fill = sentiment)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.3) + 
  geom_errorbar(aes(ymin = lower, ymax = upper)
                , position = "dodge"
                , color = "black") +
  scale_fill_manual(values = c("red2", "blue", "darkgreen")) +
  labs(title = "Syuzhet Sentiment Analysis",
       x = "Fire",
       y = "Proportion of Tweets",
       fill = "Sentiment") +
  theme_bw()
```

There are not many neutral tweets. The majority of tweets are classified as negative and the proportion of negativity appeats to be steadily increasing through time.

Figure 2 is the same graphic but with sentiment classifications from roBERTa-base model.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3, fig.width=6,fig.cap="Shift in Wildfire Sentiment Over Time."}
main %>%
  filter(followers_count > 0 & 
           following_count > 0 & 
           like_count > 0) %>%
  filter(!stringr::str_detect(description,'meteorology')) %>%
  mutate(fire = fct_relevel(fire, c("Tubbs Fire", 
                      "Camp Fire", 
                      "Bay Area Fire", 
                      "Dixie Fire"))) %>%
  group_by(fire) %>%
  summarize(Positive = length(bert_sent[bert_sent == "positive"]) / n(),
            Negative = length(bert_sent[bert_sent == "negative"]) / n(),
            Neutral = length(bert_sent[bert_sent == "neutral"]) / n(),
            Total = n()) %>%
  gather(sentiment, percentage, Positive:Neutral) %>%
  mutate(
    se = sqrt(percentage * (1 - percentage) / Total),
    lower = percentage - 2 * se,
    upper = percentage + 2 * se) %>%
  ggplot(aes(x = fire, y = percentage, fill = sentiment)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.3) + 
  geom_errorbar(aes(ymin = lower, ymax = upper)
                , position = "dodge"
                , color = "black") +
  scale_fill_manual(values = c("red2", "blue", "darkgreen")) +
  labs(title = "BERT Sentiment Analysis",
       x = "Fire",
       y = "Proportion of Tweets",
       fill = "Sentiment") +
  theme_bw()

```

The roBERTa-base model classifies a larger proportion of the tweets as neutral. It was difficult for us to remove all news related tweets. roBERTa-base model classified a lot of these as neutral when the syuzhet package would incorrectly assig the tweet as slightly positive or slightly negative. However, roBERTa-base model classified less negativity for the Tubbs fire than the syuzhet package and we see a more consistent increase in negativity, plateauing at the Bay Area Fire and Dixie Fire. It was difficult for us to remove all news related tweets, roBERTa-base model

Let's examine differences in specific tweet classifation between the syuzhet and roBERTa-base model. Specifically, for the Tubbs fire where we saw less negativity classified by the roBERTa-base model. It was difficult for us to remove all news related tweets. roBERTa-base model classified a lot of these as neutral when the syuzhet package would incorrectly assig the tweet as slightly positive or slightly negative.

**Examples of neutral for roBERTa-base and negative for syuzhet**

* Where Did the Tubbs Fire Begin? Investigators Scour the Ashes https://t.co/9z2QjBXi77

* The cause of the massive Tubbs fire is still unknown -- but officials think they may be zeroing in: https://t.co/0k9w4nFgpE.

These are examples of truly neutral tweets from news organizations. However, the syuzhet package tends to not assign many neutral tweets. These were incorrectly classified as negative by syuzhet, possibly due to words like "scour" or "unknown", but correctly seen as neutral from roBERTa-base.

**Examples of negative for roBERTa-base and positive for syuzhet**

* The personal treasures former Raiders great Cliff Branch lost in Santa Rosa fires--a truly sad tale by Bob Padecky. https://t.co/gIT0AulURq

* My parents lost everything in the Tubbs Fire yesterday. Anything will help. Please share as much as you can. Thank you so much. #tubbsfire https://t.co/xFAexYshSL

* Many Sonic staff in Santa Rosa and surrounding area are affected by the Tubbs Fire. Best wishes to all, and be safe! It is devastating. https://t.co/pVqAFisxKd

The first tweet used the word "great" to describe the football player but the overall sentiment should have been negative due to Cliff losing his possessions in the fire. Many tweets consisting of "Thank you" or "Be safe" were seen as positive from the syuzhet model, when the overall message of the tweet was in a negative context. This subtle distinction was correctly picked up by the roBERTa-base model.

**Examples of positive for roBERTa-base and negative for syuzhet**

* This Friday's midnight showing the Rocky Horror Picture Show is now a benefit for Tubbs Fire victims: https://t.co/Su7tOzcnfm

* All I can say is wow. Props to Berkeley #firefighters. Crazy video of firefighters stunned by scope of Tubbs Fire https://t.co/8kZwh6g5uy

These two examples highlight how the syuzhet package was tricked into a negative classification when the tweet should have been seen as positive. The describing of the movie has some words associated with negativity but the overall message of tweet was a benefit for Tubbs Fire victim, this was correctly picked up on by roBERTa-base. The second tweet is a user thanking the firefighters, the fact that she was stunned by the scope of the fire made syuzhet believe the tweet was negative. However, she tweeted this to give props to the firefighters. These examples highlight how the roBERTa-base model was more effective at distinguishing subtle nuances in tweets, resulting in more accurate classifications of sentiment.

### 2022-12-11 load.R

@@ -23,8 +23,8 @@ n <- 100000
 # tubbs fire: October 2017
 params = list(
   `query` = 'fire -is:retweet lang:en place_country:US',
-  `start_time` = '2020-08-15T00:00:00.00Z',
-  `end_time` = '2020-09-15T00:00:00.00Z',
+  `start_time` = '2017-10-01T00:00:00.00Z',
+  `end_time` = '2017-11-01T00:00:00.00Z',
   `max_results` = '500',
   `tweet.fields` = 'created_at,lang,public_metrics',
   `expansions` = 'geo.place_id',
@@ -78,6 +78,13 @@ while (nrow(tweets) < n) {
     simplifyDataFrame = TRUE
   )
   
+  if (!is.null(result$title)) {
+    if (result$title == 'Too Many Requests') {
+      print('Waiting for Twitter API...')
+      Sys.sleep(60)
+      next
+    } 
+  }
   
   tweet_df <- cbind(
     result$data$text,
@@ -123,8 +130,8 @@ while (nrow(tweets) < n) {
   
 }
 
-saveRDS(tweets, 'bay_area_fire_tweets.RDS')
-saveRDS(places, 'bay_area_fire_places.RDS')
+saveRDS(tweets, 'tubbs_fire_tweets.RDS')
+saveRDS(places, 'tubbs_fire_places.RDS')
 
 # t3 <- readRDS('dixie_tweets.RDS')
 

### 2022-12-11 map.R

@@ -1,3 +1,6 @@
+# This code is heavily based on Prof. David Brown's lab replicating Joshua 
+# Katz's geographic plots of lexical variants. Thanks to him for providing it!
+
 library(tidyverse)
 library(kknn)
 
@@ -12,9 +15,14 @@ require(rgeos)
 
 setwd("~/msp/wildfire_sentiment_analysis/data")
 
-tweets <- readRDS('dixie_fire_tweets.RDS')
-places <- readRDS('dixie_fire_places.RDS')
-sentiment <- read.csv('dixie_fire_sentiment.csv')
+fire_name <- 'bay_area'
+tweets <- readRDS(paste0(fire_name, '_fire_tweets.RDS'))
+places <- readRDS(paste0(fire_name, '_fire_places.RDS'))
+sentiment <- read.csv(paste0(fire_name, '_fire_sentiment.csv'))
+
+# number of words in file.
+# sum(unlist(lapply(sentiment$text, function(x) str_count(x, '\\w+'))))
+
 
 sentiment$text_cleaned <- sentiment$text
 sentiment$text <- NULL
@@ -38,10 +46,8 @@ data <- cbind(tweets, sentiment) %>%
 
 
 states_us <- rnaturalearthdata::states50
-
 states_us <- states_us[states_us$iso_a2 == 'US',]
 states_us <- states_us[ !grepl( "Alaska|Hawaii" , states_us$name ) , ]
-
 states_us <- st_as_sf(states_us)
 
 nalcc <- "+proj=lcc +lat_1=20 +lat_2=60 +lat_0=40 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
@@ -62,7 +68,7 @@ map_df <- data.frame(
   lat = st_coordinates(coord_data)[, 2]
 )
 
-
+# helpful plot to debug individual points if needed
 # ggplot(data = map_df) +
 #   geom_sf(data = us) +
 #   geom_point(aes(x = lon, y = lat, color = sentiment), alpha = 1) +
@@ -74,19 +80,16 @@ valid_split <- rsample::initial_split(map_df, .75)
 map_df_train <- rsample::analysis(valid_split)
 map_df_test <- rsample::assessment(valid_split)
 
-# knn_usa <- kknn::kknn(
-#   sentiment ~ ., 
-#   train = map_df_train, 
-#   test = map_df_test, 
-#   kernel = "gaussian" 
-#   # k = 61
-# )
-
[…]

### 2022-12-11 bert_sentiment.py
add copy of BERT code

@@ -1,0 +1,25 @@
+from transformers import AutoTokenizer, AutoModelForSequenceClassification
+from transformers import pipeline
+import pandas as pd
+import pyreadr
+import re
+
+in_df = pyreadr.read_r('C:/Users/halse/Documents/msp/wildfire_sentiment_analysis/data/tubbs_fire_tweets.RDS')
+in_df = in_df[None]
+tweets = list(in_df.text)
+# remove links
+tweets = [re.sub(r'https?://\S+', '', t) for t in tweets]
+# remove usernames
+tweets = [re.sub(r'@[^\s]+', '', t) for t in tweets]
+
+
+MODEL = f"cardiffnlp/twitter-roberta-base-sentiment-latest"
+tokenizer = AutoTokenizer.from_pretrained(MODEL)
+model = AutoModelForSequenceClassification.from_pretrained(MODEL)
+# tokenizer.save_pretrained(MODEL)
+# model.save_pretrained(MODEL)
+
+sentiment_task = pipeline('sentiment-analysis', model=model, tokenizer=tokenizer)
+out_df = pd.DataFrame(sentiment_task(tweets))
+out_df['text'] = tweets
+out_df.to_csv('tubbs_fire_sentiment.csv', index=False)

### 2022-12-12 paper.Rmd
Updated Introduction

@@
 
+## Abstract
+
 ## Introduction
@@
 According to the National Interagency Fire Center (NIFC), more acreage burned nationally in each of the last seven years than any single year but one in the first ten years this data was collected (1983-1993, ex.
-1988).^[See https://www.nifc.gov/fire-information/statistics/wildfires.] We sought to analyze if and how public sentiment has shifted over time with respect to these cataclysmic events, with the hopes of gaining some insight into public perceptions of climate change.
+1988).^[See https://www.nifc.gov/fire-information/statistics/wildfires.] With more and more people affected by the increase in wildfire activity, it's reasonable to infer an increasing outcry of negativity towards this devastating symptom of climate change.
+However, it is hypothesized that the negative outcry may only come from people directly affected by the fire.
+An, "out of sight, out of mind", mentality may be prevalent in the United States, particularly around the affects of wildfires.
+We explored four recent US wildfires, that took place sequentially with the past 5 years, to analyze if and how public sentiment has shifted over time with respect to these cataclysmic events.
+Specifically, we sought to answer the following two questions.
+
+* How has public sentiment towards wildfires changed in the past 5 years?
+
+* Is this change regionalized? Is public sentiment only shifted by those directly affected by these devastating events?
+
+We compared two different sentiment analysis algorithms in hopes of gaining some insight into public perceptions of climate change.
 
@@
 
+# Sentiment Analysis
+
 We used two sentiment analysis algorithms of varying complexity to classify the aforementioned tweets.
@@
 
+# Change through time
+
 To track to sentiment shift through time, we chronologically ordered the fires (Tubbs Fire, Camp Fire, Bay Area Fire, Dixie Fire) and calculated the proportion of each sentiment for each fire.
 We used proportions because of the unequal distribution of tweets across the fires.
+
+# Effect of location
 
@@
 ## Results
+
+# Change through time
 
@@
 It was difficult for us to remove all news related tweets, roBERTa model
+
+# Effect of Location
+
+# Comparison of Sentiment Analysis Algorithm
 
@@
 
+[Add subheaders for each section]
+
 Even though the tweet volume was not consistent throughout time, Figure 1 gives us preliminary evidence that there may have been a shift in tweet sentiment towards wildfires in the last five years.

### 2022-12-12 paper.Rmd
final(ish) draft

--- 2022-12-11 paper.Rmd
@@
 ---
-title: "INSERT"
+title: "Trial by Fire: Diachronic and Regional Sentiment Towards American Wildfires"
 author: "Charles Hutchinson, Sam Kalman"
@@
 
+## Abstract
+
+We analyzed diachronic regional trends in public tweets about wildfires to better understand public perception towards climate change.
+Processing over 350,000 tweets, we used both a dictionary-based approach in the `syuzhet` R package and a roBERTa model from researchers at Cardiff University for sentiment analysis.
+We found that sentiment has become increasingly negative over time, particularly in the Pacific Northwest.
+This conclusion could be reinforced by directly comparing air quality and sentiment.
+
 ## Introduction
 
-One of North America's most noticeable symptoms of climate change has been a dramatic increase in the volume and intensity of wildfires, particularly in its western regions.
+One of North America's most noticeable symptoms of climate change has been a dramatic increase in the volume and intensity of wildfires in its western regions.
 August and September were previously defined by clear skies and fair weather, but fire and smoke have become a recurrent factor in this area.
 According to the National Interagency Fire Center (NIFC), more acreage burned nationally in each of the last seven years than any single year but one in the first ten years this data was collected (1983-1993, ex.
-1988).^[See https://www.nifc.gov/fire-information/statistics/wildfires.] We sought to analyze if and how public sentiment has shifted over time with respect to these cataclysmic events, with the hopes of gaining some insight into public perceptions of climate change.
+1988).^[See https://www.nifc.gov/fire-information/statistics/wildfires.] With more and more people affected by the increase in wildfire activity, it's reasonable to infer an increasing outcry of negativity towards this devastating symptom of climate change.
+However, less is known about regional attitudes towards this recent trend.
+As with many other catastrophes, is it true that wildfires are "out of sight, out of mind" for those not directly affected?
+
+We explored four recent US wildfires that took place in the past five years to analyze if and how public sentiment has shifted over time with respect to these cataclysmic events.
+Specifically, we sought to answer the following two questions:
+
+1. How has public sentiment towards wildfires changed in the past five years?
+2. Is this change regionalized? Are changes in sentiment (if any) limited to those directly affected by these devastating events?
+
+We compared two different sentiment analysis algorithms in the hopes of gaining some insight into changing public perceptions of climate change.
 
@@
 ) %>%
-  kableExtra::kbl(caption = "Relevant Tweets per Recent Wildfire.", booktabs = T, linesep = "") %>%
+  kableExtra::kbl(caption = "Tweet Counts by Wildfire.", booktabs = T, linesep = "") %>%
   kableExtra::kable_styling(latex_options = "HOLD_position") %>%
@@
 This proved to be particularly debilitating when working on location-based analyses, since only a subset of that limited population reported their location while tweeting.
-To alleviate this problem, we chose to use much more generic search terms.
+To alleviate this problem, we chose to use much more generic search terms when conducting our location-based analyses.
 For the purposes of this paper, we found the search term "fire" tended to return the most relevant results (as compared to similarly broad terms like "smoke" or "burn"), though this was not a systematic decision.
@@
 
-We continued to use the same date windows for each fire.
+In our location-based analyses, we continued to use the same date windows for each fire.
 As before, we only pulled English tweets, and filtered out any retweets.
-We asked the Twitter API to filter for tweets in the United States, then manually filtered those tweets down to those made in the contiguous lower 48 with longitude and latitude filters.
+We asked the Twitter API to filter for tweets in the United States, then manually filtered those tweets down to those made in the continental US with longitude and latitude filters.
 Twitter provides locations in a bounding box, not on a single point, so we assumed that each tweet was made in the center of the bounding box.
@@
 ) %>%
-  kableExtra::kbl(caption = "Location-Enabled Tweet Counts", booktabs=T, linesep = "") %>%
+  kableExtra::kbl(caption = "Location-Enabled Tweet Counts by Wildfire.", booktabs=T, linesep = "") %>%
   kableExtra::kable_styling(latex_options = "HOLD_position") %>%
@@
 ```
+
+We found that most tweets from news organization contained neutral, non-opinionated information, which wasn’t useful for our research question.
+As such, any tweet made by a user with the word "meteorology" in their Twitter biography was not considered in our analysis.
+Additionally, in an attempt to filter out bots adding unnecessary noise to the data, we only used tweets from users with at least one follower, and the tweet itself needed at least one external interaction (a like/retweet/comment).
+It's worth noting that the authors of Cardiff's roBERTa model (Loureriro et al., 2022) addressed the bot issue by filtering out the top 1% of users by volume, but we did not have a similarly unbiased data set to determine an appropriate posting cutoff.
 
@@
 The first algorithm we used was from the `syuzhet` package^[https://www.rdocumentation.org/packages/syuzhet/versions/1.0.6] in R.
-The syuzhet package returns sentiment scores on a spectrum of positively/negativity using a sentiment dictionary developed in the Nebraska Literacy Labs.
+The `syuzhet` package returns sentiment scores on a spectrum of positively/negativity using a sentiment dictionary developed in the Nebraska Literacy Labs.
 We deemed any sentiment score that was greater than zero to be positive, less than zero, negative, and exactly equal to zero, neutral.
-Unfortunately, the `syuzhet` package wasn't designed or trained on short, declarative statements such as tweets.
+Unfortunately, the `syuzhet` package wasn't designed for or trained on short, declarative statements such as tweets.
 
-To overcome this issue, we decided to use Cardiff's roBERTa-base model for sentiment analysis.^[https://huggingface.co/cardiffnlp/twitter-roBERTa-sentiment-latest] This model was trained on roughly 124 million tweets from January 2018 to December 2021 and used the TweetEval benchmark for fine tuning the sentiment analysis.
-The algorithm outputs probabilities that a tweet is positive, neutral, or negative.
+To overcome this issue, we decided to use Cardiff's roBERTa-base model for sentiment analysis as well.^[https://huggingface.co/cardiffnlp/twitter-roBERTa-sentiment-latest] This model was trained on roughly 124 million tweets from January 2018 to December 2021, using the TweetEval benchmark to fine tune the sentiment analysis.
+The model outputs probabilities that a tweet is positive, neutral, or negative.
 For the purposes of our analysis, we simplified the output to a categorical variable representing the sentiment with the highest probability.
@@
 
-We found that most tweets from news organization contained neutral, non-opinionated information, which wasn’t useful for our research question.
-Therefore ,any tweet that came from a user with the word "meteorology" in their bio was not considered in our analysis.
-Additionally, in an attempt to filter out bots adding unnecessary noise to the data, we only used tweets from users with at least one follower, and the tweet itself needed at least one external interaction (a like/retweet/comment).
-This was done consistently before running both of the discussed sentiment algorithms.
-It's worth noting that the authors of Cardiff's roBERTa model (Loureriro et al., 2022) address the bot issue by filtering out the top 1% of users by volume, but we did not have a similarly unbiased data set to determine an appropriate posting cutoff.
-
 To track to sentiment shift through time, we chronologically ordered the fires (Tubbs Fire, Camp Fire, Bay Area Fire, Dixie Fire) and calculated the proportion of each sentiment for each fire.
@@
 
-We first converted from longitude/latitude coordinates to WGS84 coordinates, then projected these coordinates using a Lambert Conformal Conic projection (LCC).
+Plotting sentiment by location proved to be a more involved task.
+As stated in the Data section, we chose to simplify the provided bounding boxes for each tweet to the center point.
+From there, we converted from longitude/latitude coordinates to WGS84 coordinates, then projected these coordinates using a Lambert Conformal Conic projection (LCC).
 This process was completed using the `sf` package.
-[ADD MORE HERE]
+Using these transformed coordinates and the corresponding sentiments, we used weighted k-nearest neighbors with a Gaussian kernel to interpolate probabilities of a given sentiment across the map.
 
 ## Results
+
+### Change Over Time
 
@@
 The error bars reflect a 95% confidence interval for each proportion.
+There are not many neutral tweets, though our definition of neutrality with the `syuzhet` scoring is quite stringent.
+The majority of tweets are classified as negative, and the proportion of negativity increases over time.
 
-```{r, fig.height=3, fig.width=6,fig.cap="Shift in Wildfire Sentiment Over Time."}
+```{r, fig.height=3, fig.width=6, fig.cap="Shift in Wildfire Sentiment Over Time."}
 
@@
 
-There are not many neutral tweets, though our definition of neutrality with the `syuzhet` scoring is quite stringent.
-The majority of tweets are classified as negative, and the proportion of negativity appears to be increasing through time.
-
-We repeat the same analysis using scores from Cardiff's roBERTa model in Figure 2.
-
 ```{r, fig.height=3, fig.width=6,fig.cap="Shift in Wildfire Sentiment Over Time."}
@@
 
+We repeat the same analysis using scores from Cardiff's roBERTa model in Figure 2.
 The roBERTa model classifies a much larger proportion of the tweets as neutral.
@@
 We found that this iteration was very effective at putting "inorganic" reactions from news organizations in this neutral category.
-However, the roBERTa model classified less negativity for the Tubbs fire than the syuzhet package and we see a more consistent increase in negativity, plateauing at the Bay Area Fire and Dixie Fire.
-It was difficult for us to remove all news related tweets, roBERTa model
+Additionally, the roBERTa model classified less negativity for the Tubbs fire than the `syuzhet` package and we see a more consistent increase in negativity, plateauing at the Bay Area Fire and Dixie Fire.
 
-Let's examine differences in specific tweet classifation between the syuzhet and roBERTa model.
-Specifically, for the Tubbs fire where we saw less negativity classified by the roBERTa model.
-It was difficult for us to remove all news related tweets.
-roBERTa model classified a lot of these as neutral when the syuzhet package would incorrectly assig the tweet as slightly positive or slightly negative.
+### Effect of Location
 
-**Examples of neutral for roBERTa and negative for syuzhet**
+Figure 3 shows the interpolated roBERTa sentiment probabilities across the continental US.
+The obvious trend is one of consolidation - in particular, the Pacific Northwest has become increasingly negative about fires during fire season.
+As one of the hardest hit regions by these catastrophes in the country, this is not a surprising result.
 
-* Where Did the Tubbs Fire Begin? Investigators Scour the Ashes https://t.co/9z2QjBXi77
+It is important to reiterate that these maps use a more generic data set than the bar plots above.
+Specific mention of the fire's name was not required - only the term "fire" was needed.
+There are a wide variety of ways that the term "fire" is used on Twitter, many of which are not related to wildfires.
+Manual inspection revealed that this accounts for the large swaths of positive tweets in plots such as Figures 3a and 3d.
+However, we were unable to find a robust explanation for the regional variation in positivity over time.
 
-* The cause of the massive Tubbs fire is still unknown -- but officials think they may be zeroing in: https://t.co/0k9w4nFgpE.
+The intensity of color in Figure 3 corresponds to probability of a given sentiment, not volume.
+Plotting by the latter would result in a mostly blank map, with a few spots of color around urban centers.
 
-These are examples of truly neutral tweets from news organizations.
-However, the syuzhet package tends to not assign many neutral tweets.
-These were incorrectly classified as negative by syuzhet, possibly due to words like "scour" or "unknown", but correctly seen as neutral from roBERTa.
-
-**Examples of negative for roBERTa and positive for syuzhet**
-
-* The personal treasures former Raiders great Cliff Branch lost in Santa Rosa fires--a truly sad tale by Bob Padecky. https://t.co/gIT0AulURq
-
-* My parents lost everything in the Tubbs Fire yesterday. Anything will help. Please share as much as you can. Thank you so much. #tubbsfire https://t.co/xFAexYshSL
-
-* Many Sonic staff in Santa Rosa and surrounding area are affected by the Tubbs Fire. Best wishes to all, and be safe! It is devastating. https://t.co/pVqAFisxKd
-
-The first tweet used the word "great" to describe the football player but the overall sentiment should have been negative due to Cliff losing his possessions in the fire.
-Many tweets consisting of "Thank you" or "Be safe" were seen as positive from the syuzhet model, when the overall message of the tweet was in a negative context.
-This subtle distinction was correctly picked up by the roBERTa model.
-
-**Examples of positive for roBERTa and negative for syuzhet**
-
-* This Friday's midnight showing the Rocky Horror Picture Show is now a benefit for Tubbs Fire victims: https://t.co/Su7tOzcnfm
-
-* All I can say is wow. Props to Berkeley #firefighters. Crazy video of firefighters stunned by scope of Tubbs Fire https://t.co/8kZwh6g5uy
-
-These two examples highlight how the syuzhet package was tricked into a negative classification when the tweet should have been seen as positive.
-The describing of the movie has some words associated with negativity but the overall message of tweet was a benefit for Tubbs Fire victim, this was correctly picked up on by roBERTa.
-The second tweet is a user thanking the firefighters, the fact that she was stunned by the scope of the fire made syuzhet believe the tweet was negative.
-However, she tweeted this to give props to the firefighters.
-These examples highlight how the roBERTa model was more effective at distinguishing subtle nuances in tweets, resulting in more accurate classifications of sentiment.
-
-Important to note that Tubbs was much shorter, and the other summer fires were usually coincident with a broader wildfire season. [ADD MORE HERE]
-
-```{r, fig.cap = "Tweet Sentiment (roBERTa) by Location", fig.height = 7, fig.subcap = c("Dixie (2021)", "Bay Area (2020)", "Camp (2018)", "Tubbs (2017)"), fig.ncol = 2, out.width = "50%", fig.align='center'}
+```{r, fig.cap = "Tweet Sentiment (roBERTa) by Location.", fig.height = 7, fig.subcap = c("Dixie (2021)", "Bay Area (2020)", "Camp (2018)", "Tubbs (2017)"), fig.ncol = 2, out.width = "50%", fig.align='center'}
 
@@
 
+### Comparison of Models
+
+While performing manual checks of each model's sentiment classification, we were consistently impressed by the performance of Cardiff's model.
+The BERT framework was able to capture context in a way that was impossible for `syuzhet`'s dictionary-based approach.
+We elaborate on some common cases we saw below.
+
+**1. Neutral roBERTa + Negative `syuzhet`**
+
+> Where Did the Tubbs Fire Begin? Investigators Scour the Ashes https://t.co/9z2QjBXi77
+
+> The cause of the massive Tubbs fire is still unknown -- but officials think they may be zeroing in: https://t.co/0k9w4nFgpE.
+
+These are examples of neutral tweets from news organizations.
+These were incorrectly classified as negative by `syuzhet`, due to words like "scour" or "unknown", though we note that the requirement of a zero sentiment score for neutrality contributed to this incorrect classification.
+
+**2. Positive roBERTa + Negative `syuzhet`**
+
+> This Friday's midnight showing the Rocky Horror Picture Show is now a benefit for Tubbs Fire victims: https://t.co/Su7tOzcnfm
+
+> All I can say is wow. Props to Berkeley #firefighters. Crazy video of firefighters stunned by scope of Tubbs Fire https://t.co/8kZwh6g5uy
+
+In the first tweet, `syuzhet` is misled by the title of the movie (Rocky Horror Picture Show), while roBERTa identifies the altruistic context (a charity showing).
+In the second tweet, `syuzhet` overweights the negative language with respect to the fire, while roBERTa correctly captures the heroic context of the tweet.
+These examples highlight how the roBERTa model was more effective at distinguishing subtle nuances in tweets, resulting in more accurate classifications of sentiment.
+
+**3. Negative roBERTa + Positive `syuzhet`**
+
+> The personal treasures former Raiders great Cliff Branch lost in Santa Rosa fires--a truly sad tale by Bob Padecky. https://t.co/gIT0AulURq
+
+> My parents lost everything in the Tubbs Fire yesterday. Anything will help. Please share as much as you can. Thank you so much. #tubbsfire https://t.co/xFAexYshSL
+
+> Many Sonic staff in Santa Rosa and surrounding area are affected by the Tubbs Fire. Best wishes to all, and be safe! It is devastating. https://t.co/pVqAFisxKd
+
+The first tweet uses the word "great" to describe the football player Cliff Branch, but the overall sentiment should be negative due to Branch losing his possessions in the fire.
+Many tweets containing the terms "thank you" or "be safe" were seen as positive by the `syuzhet` model, when the overall message of the tweet was in a negative context.
+For the purposes of our analysis, this distinction was the most important improvement between roBERTa and `syuzhet`'s output.
+
 ## Discussion
 
-Even though the tweet volume was not consistent throughout time, Figure 1 gives us preliminary evidence that there may have been a shift in tweet sentiment towards wildfires in the last five years.
-Beginning with Camp Fire, the amount of negative tweets significantly increased for each wildfire.
+Figures 1 and 2 shows evidence of a negative shift in sentiment towards wildfires in the last five years.
+The increase in negativity is broadly consistent between sentiment models.
+Figure 3 shows an increasing concentration of negative sentiment towards fire in the Pacific Northwest.
+We propose that negative sentiment towards wildfires has become increasingly localized over the past five years.
 
@@
 
-It was surprising that we could not find more tweets for the Bay Area and Tubbs fires, given their proximity to major urban centers.
+In our first data set (Table 1), sample size varied by more than an order of magnitude between certain fires.
+It was surprising that we could not find more tweets for the Bay Area and Tubbs fires, given their proximity to San Francisco.
 We originally suspected that the Camp Fire tweet set contained many false positives, but it appears that fire generated more engagement than expected due to the almost total devastation (and subsequent press coverage) of the towns of Paradise, Concow, Magalia, and Butte Creek Canyon.
@@
 As the largest ever single wildfire in the state's history, part of its impact included popular Lassen National Park, the coverage of which likely increased engagement on Twitter.
+
+Given more time, it would have been helpful to correct for fire intensity.
+These fires were all record-setting, but they took place in different conditions.
+Some were closer to urban centers, and some were quite far.
+Some took place in the middle of summer and were coincident with a large set of other wildfires, while others (particularly the Camp Fire) took place later in the season, in near isolation.
+Early September of 2020, near the peak of the Bay Area fires, was a particularly striking example of how catastrophic these coincident fires could be (see Figure 4).
+It would be particularly interesting to compare sentiment to air quality, though we suspect the combination of (typically) rapid fluctuations in air quality and the sparsity of location-enabled tweets on any given day would make robust statistical conclusions difficult.
+
+ ![Wildfire Smoke as of September 9, 2020. By [NASA Earth Observatory](https://earthobservatory.nasa.gov/images/147261/a-wall-of-smoke-on-the-us-west-coast?src=nha).](Western_fires_2020.jpg)

### 2022-12-12 map.R

@@ -15,7 +15,7 @@ require(rgeos)
 
 setwd("~/msp/wildfire_sentiment_analysis/data")
 
-fire_name <- 'bay_area'
+fire_name <- 'dixie'
 tweets <- readRDS(paste0(fire_name, '_fire_tweets.RDS'))
 places <- readRDS(paste0(fire_name, '_fire_places.RDS'))
 sentiment <- read.csv(paste0(fire_name, '_fire_sentiment.csv'))

### 2022-12-12 paper.Rmd
minor edits

@@
 data.frame(
-  Fire = c('Dixie', 'Bay Area', 'Camp', 'Tubbs'),
+  Fire = c('Dixie Fire', 'Bay Area Fires', 'Camp Fire', 'Tubbs Fire'),
   Tweets = c(36353, 71608, 84284, 76839),
@@
 ) %>%
+  arrange(desc(row_number())) %>%
   kableExtra::kbl(caption = "Location-Enabled Tweet Counts by Wildfire.", booktabs=T, linesep = "") %>%
@@
 
-To track to sentiment shift through time, we chronologically ordered the fires (Tubbs Fire, Camp Fire, Bay Area Fire, Dixie Fire) and calculated the proportion of each sentiment for each fire.
+To track to sentiment shift through time, we chronologically ordered the fires (Tubbs Fire, Camp Fire, Bay Area Fires, Dixie Fire) and calculated the proportion of each sentiment for each fire.
 We used proportions because of the unequal distribution of tweets across the fires.
+We apply the Bonferroni correction to our 95% confidence intervals to account for multiple testing.
 
@@
 In Figure 1, we plot the sentiment classification proportions for each of the four wildfires.
-The fires are in chronological order, beginning with Tubbs Fire and ending with the Dixie Fire.
+The fires are in chronological order, beginning with the Tubbs Fire and ending with the Dixie Fire.
 The error bars reflect a 95% confidence interval for each proportion.
@@
     se = sqrt(percentage * (1 - percentage) / Total),
-    lower = percentage - 2 * se,
-    upper = percentage + 2 * se) %>%
+    lower = percentage + qnorm(0.05*2/choose(3, 2))*se,
+    upper = percentage + qnorm(1 - 0.05*2/choose(3, 2))*se) %>%
   ggplot(aes(x = fire, y = percentage, fill = sentiment)) +
@@
 We found that this iteration was very effective at putting "inorganic" reactions from news organizations in this neutral category.
-Additionally, the roBERTa model classified less negativity for the Tubbs fire than the `syuzhet` package and we see a more consistent increase in negativity, plateauing at the Bay Area Fire and Dixie Fire.
+Additionally, the roBERTa model classified less negativity for the Tubbs fire than the `syuzhet` package and we see a more consistent increase in negativity, plateauing at the Bay Area Fires and Dixie Fire.
 

### 2022-12-12 paper.Rmd
Updated References

@@
 
-To overcome this issue, we decided to use Cardiff's roBERTa-base model for sentiment analysis as well.^[https://huggingface.co/cardiffnlp/twitter-roBERTa-sentiment-latest] This model was trained on roughly 124 million tweets from January 2018 to December 2021, using the TweetEval benchmark to fine tune the sentiment analysis.
+To overcome this issue, we decided to use Cardiff's roBERTa-base model for sentiment analysis as well.^[https://huggingface.co/cardiffnlp/twitter-roberta-base-sentiment-latest] This model was trained on roughly 124 million tweets from January 2018 to December 2021, using the TweetEval benchmark to fine tune the sentiment analysis.
 The model outputs probabilities that a tweet is positive, neutral, or negative.
@@
 
+## References
+
+Anon. n.d. “Wildfires and Acres.” Nifc.gov. \
+(https://www.nifc.gov/fire-information/statistics/wildfires)
+
+Jockers, M. (2020), “Syuzhet Package - RDocumentation.” Rdocumentation.org. \
+(https://www.rdocumentation.org/packages/syuzhet/versions/1.0.6).
+
+Lou, D & Camacho-Collados, J (2022).
+“Cardiffnlp/Twitter-Roberta-Base-Sentiment-Latest · Hugging Face.” 
+Huggingface.Co. \
+(https://huggingface.co/cardiffnlp/twitter-roberta-base-sentiment-latest).
+
+Sagar, Chaitanya. 2018. “Twitter Sentiment Analysis Using R.” Dataaspirant. \ 
+(https://dataaspirant.com/twitter-sentiment-analysis-using-r/).
+
+### Please see submitted GitHub Repository for associated Code Appendix
+
  ![Wildfire Smoke as of September 9, 2020. By [NASA Earth Observatory](https://earthobservatory.nasa.gov/images/147261/a-wall-of-smoke-on-the-us-west-coast?src=nha).](Western_fires_2020.jpg)

### 2022-12-12 paper.Rmd
minor edits

@@
 ---
-title: "Trial by Fire: Diachronic and Regional Sentiment Towards American Wildfires"
+title: "Trial by Fire: Diachronic Regional Sentiment Towards American Wildfires"
 author: "Charles Hutchinson, Sam Kalman"
@@
 
+![Wildfire Smoke as of September 9, 2020. By [NASA Earth Observatory](https://earthobservatory.nasa.gov/images/147261/a-wall-of-smoke-on-the-us-west-coast?src=nha).](Western_fires_2020.jpg)
+
 ## References
@@
 
-Lou, D & Camacho-Collados, J (2022).
+Loureiro, D. et al. (2022).
 “Cardiffnlp/Twitter-Roberta-Base-Sentiment-Latest · Hugging Face.” 
@@
 
-### Please see submitted GitHub Repository for associated Code Appendix
+## Code Appendix
 
- ![Wildfire Smoke as of September 9, 2020. By [NASA Earth Observatory](https://earthobservatory.nasa.gov/images/147261/a-wall-of-smoke-on-the-us-west-coast?src=nha).](Western_fires_2020.jpg)
+Please find all code for this analysis at https://github.com/halseyhutch/wildfire_sentiment_analysis.

### 2022-12-12 paper.Rmd
final draft

@@
 One of North America's most noticeable symptoms of climate change has been a dramatic increase in the volume and intensity of wildfires in its western regions.
-August and September were previously defined by clear skies and fair weather, but fire and smoke have become a recurrent factor in this area.
-According to the National Interagency Fire Center (NIFC), more acreage burned nationally in each of the last seven years than any single year but one in the first ten years this data was collected (1983-1993, ex.
-1988).^[See https://www.nifc.gov/fire-information/statistics/wildfires.] With more and more people affected by the increase in wildfire activity, it's reasonable to infer an increasing outcry of negativity towards this devastating symptom of climate change.
+August and September were previously defined by clear skies and fair weather, but fire and smoke have become recurrent factors in this area.
+Since 2015, more acreage has burned nationally each year compared to any year between 1983 (the first year this data was collected) and 1993, excluding 1988.^[See https://www.nifc.gov/fire-information/statistics/wildfires.] With more and more people affected by this increase in wildfire activity, it's reasonable to infer an increasing outcry of negativity towards this symptom of climate change.
 However, less is known about regional attitudes towards this recent trend.
@@
 2. Is this change regionalized? Are changes in sentiment (if any) limited to those directly affected by these devastating events?
-
-We compared two different sentiment analysis algorithms in the hopes of gaining some insight into changing public perceptions of climate change.
 
@@
 
-The size of each fire's corpus was unfortunately still far from uniform, possibly due to unequal media coverage, but we found this to be a reasonable compromise among a set of unappealing options.
-We found both Camp Fire and Dixie Fire generated far more content on Twitter under our listed constraints when compared to Tubbs Fire and the Bay Area Fires.
+The size of each fire's corpus was unfortunately still far from uniform, possibly due to unequal media coverage, but we found this to be a reasonable compromise among a set of imperfect options.
+We found that both the Camp Fire and the Dixie Fire generated far more content on Twitter under our listed constraints when compared to the Tubbs Fire and the Bay Area Fires.
 Table 1 shows the volume of tweets per fire.
@@
 Filtering for tweets that explicitly mentioned the fire by name significantly limited the total number of results.
-This proved to be particularly debilitating when working on location-based analyses, since only a subset of that limited population reported their location while tweeting.
+This proved to be particularly restricting when working on location-based analyses, since only a subset of that limited population reported their location while tweeting.
 To alleviate this problem, we chose to use much more generic search terms when conducting our location-based analyses.
@@
 
-To overcome this issue, we decided to use Cardiff's roBERTa-base model for sentiment analysis as well.^[https://huggingface.co/cardiffnlp/twitter-roberta-base-sentiment-latest] This model was trained on roughly 124 million tweets from January 2018 to December 2021, using the TweetEval benchmark to fine tune the sentiment analysis.
+To overcome this issue, we decided to additionally use Cardiff's roBERTa-base model for sentiment analysis.^[https://huggingface.co/cardiffnlp/twitter-roberta-base-sentiment-latest] This model was trained on roughly 124 million tweets from January 2018 to December 2021, using the TweetEval benchmark to fine tune the sentiment analysis.
 The model outputs probabilities that a tweet is positive, neutral, or negative.
@@
 
-To track to sentiment shift through time, we chronologically ordered the fires (Tubbs Fire, Camp Fire, Bay Area Fires, Dixie Fire) and calculated the proportion of each sentiment for each fire.
+To track changes in sentiment over time, we chronologically ordered the fires (Tubbs Fire, Camp Fire, Bay Area Fires, Dixie Fire) and calculated the proportion of each sentiment for each fire.
 We used proportions because of the unequal distribution of tweets across the fires.
-We apply the Bonferroni correction to our 95% confidence intervals to account for multiple testing.
+Additionally, we applied the Bonferroni correction to our 95% confidence intervals to account for multiple testing.
 
@@
 The Dixie Fire was similarly destructive.
-As the largest ever single wildfire in the state's history, part of its impact included popular Lassen National Park, the coverage of which likely increased engagement on Twitter.
+As the largest ever single wildfire in the state's history, part of its impact included popular Lassen National Park, the coverage of which may have increased engagement on Twitter.
 
@@
 Some took place in the middle of summer and were coincident with a large set of other wildfires, while others (particularly the Camp Fire) took place later in the season, in near isolation.
-Early September of 2020, near the peak of the Bay Area fires, was a particularly striking example of how catastrophic these coincident fires could be (see Figure 4).
+Satellite imagery taken near the peak of the Bay Area fires illustrates the extent to which the West Coast was engulfed in smoke, much of which was from other, smaller fires (see Figure 4).
 It would be particularly interesting to compare sentiment to air quality, though we suspect the combination of (typically) rapid fluctuations in air quality and the sparsity of location-enabled tweets on any given day would make robust statistical conclusions difficult.
